---
title: "First Draft of Tejon Analysis"
author: "sbsambado"
date: "9/3/2020"
output: html_document
---

Data were collected by the Young lab at UCSB.

Data come from Tejon Ranch in southern California. 

Experimental design:
  3 Sites that vary in elevation
    Arid (lowest), Intermediate (middle), Mesic (highest)
      block desing: 1, 2, 3
  
  Within each site there are 3 Plot types that vary in animal exclusion
    Open (no exclusion), Partial (excludes non-native larger animals), Total (only smaller animals)
    

  
  
Data include:
  Tick drags 
    across : Years (2016 - 2019) and months
    species : I. pacificus, D. occidentalis, D. variabiles, Unknown
    total ticks
  
  Mammal trapping
    across : Years (2018, 2019) and months
    species : peromyscus, and limited non-peromysucs 
    

Notes : 
- if < 1 animal for species, it was placed in the NA category (i.e. TEBO) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plyr)
library(car)
library(multcomp)
library(multcompView)
library(reshape2)
library(stringr)
library(car)
library(psych)
library(knitr)
library(nlme)
library(lme4)
library(kableExtra)
library(MASS)
library(grid)
library(vegan)
library(devtools)
library(ggfortify)
library(jtools)
devtools::install_github('gavinsimpson/ggvegan')
library(tadaatoolbox)

clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('red4','red2','red')

```

Tick Analysis
- imported data from sheet called 'tejon_all_tick_drags_09022020.csv'
```{r data importing and organizing}
# full dataset
alldrag <- read.csv("~/Desktop/tejon_all_tick_drags_09022020.csv")

# break up by site, no other filter
intermediate <- filter(alldrag, Site == 'Intermediate')
mesic <- filter(alldrag, Site == 'Mesic')
arid <- filter(alldrag, Site == 'Arid')


# aggregate tick count data
aggregateticks <- alldrag %>%
  group_by(Site) %>%
  group_by(Month) 

aggregate <-aggregate(aggregateticks$Total_ticks, 
                by = list(Month = aggregateticks$Month, 
                          Site = aggregateticks$Site, 
                          Plot = aggregateticks$Plot), FUN = sum)
colnames(aggregate) <- c('Month', 'Site','Plot','Total_ticks')
print(aggregate)


# aggregate tick count data by year
aggregateyear <-aggregate(aggregateticks$Total_ticks, 
                by = list(Month = aggregateticks$Month, 
                          Site = aggregateticks$Site, 
                          Plot = aggregateticks$Plot,
                          Year = aggregateticks$Year), FUN = sum)
colnames(aggregateyear) <- c('Month', 'Site','Plot','Year', 'Total_ticks')


# transform variables into specific types of data
alldrag$Month <- factor(alldrag$Month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))

alldrag$Site <- factor(alldrag$Site,
                       levels = c('Arid', 'Intermediate', 'Mesic'))

alldrag$Plot <- factor(alldrag$Plot,
                       levels = c('Open', 'Partial', 'Total'))

alldrag$Year <- factor(alldrag$Year,
                       levels = c('2016', '2017', '2018', '2019'))
```


Plots
one = Total ticks per Sampling by  Site no outliers 
two = Total ticks per Sampling by  Site with outliers

Does site type (X) impact total tick abundance (Y)?
```{r total tick abundance per sampling ~ site}

## represents average total ticks collected across the years and sitess

## no outliers
one <- ggplot(data = aggregate, aes(x = Site, y = Total_ticks))+
  geom_boxplot(fill = sitecolor, outlier.color = 'black', outlier.shape = NA)+
  clean_background+
  coord_cartesian(ylim = c(0, 200))+
  xlab('Site Types') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Site \n no outliers')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold'))

## with outliers
two <- ggplot(data = aggregate, aes(x = Site, y = Total_ticks))+
  geom_boxplot(fill = sitecolor, outlier.color = 'black', outlier.shape = 1)+
  clean_background+
  xlab('Site Types') + ylab('Total Ticks')+
  ggtitle('Total ticks by Site')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold')) 


one
two




```

Plots
three = Total ticks per Sampling by  Plot no outliers 
four = Total ticks per Sampling by  Plot with outliers


Does plot type (X) impact total tick abundance (Y)?
```{r total tick abundance per sampling ~ plot}

## represents average total ticks collected across the years and plots

## no outliers
three <- ggplot(data = aggregate, aes(x = Plot, y = Total_ticks))+
  geom_boxplot(fill = plotcolor, outlier.color = 'black', outlier.shape = NA)+
  clean_background+
  coord_cartesian(ylim = c(0,130))+
  xlab('Plot Types') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Plot \n no outliers')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold'))

## with outliers
four <- ggplot(data = aggregate, aes(x = Plot, y = Total_ticks))+
  geom_boxplot(fill = plotcolor, outlier.color = 'black', outlier.shape = 1)+
  clean_background+
  xlab('Plot Types') + ylab('Total Ticks')+
  ggtitle('Total ticks by Plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold')) 


three
four

```


Plots
five = Total ticks per sampling across month by site Arid (no outliers)
six = Total ticks per sampling across month by site Intermediate (no outliers)
seven = Total ticks per sampling across month by site Mesic (no outliers)
eight = compares total ticks per sampling across month by all sites (this is messy, but wanted a quick side by side picture)
nine = Total ticks per sampling across month by site Arid (no outliers)
ten = Total ticks per sampling across month by site Intermediate (no outliers)
eleven = Total ticks per sampling across month by site Mesic (no outliers)

How does total tick abundance vary by month at each site? 

```{r total tick abundance across month by site using aggregate data}
## no outliers
five <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = NA)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,13))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Arid : Total ticks per sampling by plot \n no outliers')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))

six <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = NA)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,40))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Intermediate : Total ticks per sampling by plot \n no outliers')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))

seven <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = NA)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,20))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Mesic : Total ticks per sampling by plot \n no outliers')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))

## try to arrange in a visual way
eight<- ggarrange(five +rremove('xlab') +rremove('ylab') + rremove('legend'),
                  six +rremove('xlab') +rremove('ylab') + rremove('legend'), 
                  seven +rremove('xlab') +rremove('ylab')+ rremove('legend'),
                  ncol = 3)
annotate_figure(eight,
                top = text_grob('Total Ticks by Month',
                                face = 'bold', size = 16,
                                color = 'red'))

## Outliers
nine <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = 1)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,45))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Arid : Total ticks per sampling by plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))

ten <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = 1)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,50))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Intermediate : Total ticks per sampling by plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))

eleven <- ggplot(data = aggregate, 
               aes(x = Month, y = Total_ticks,
                   fill = Plot), group = Plot)+
  geom_boxplot(outlier.shape = 1)+
  scale_color_identity()+
  facet_wrap(~Plot, ncol = 3)+
  clean_background +
  #coord_cartesian(ylim = c(0,20))+
  xlab('Months') + ylab('Total ticks per sampling ')+
  ggtitle('Mesic : Total ticks per sampling by plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(face = 'bold', size = 8)) +
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12))


five
six
seven
nine
ten
eleven
```


How does tick abundance vary across the months by Site?
```{r aggregate data}

# I know there is a better way/shorter code of doing this but it is what it is

### summaries by month

Aridmonthdrag16 <- alldrag %>%
  filter(Site == 'Arid') %>%
  group_by(Month) %>%
  filter(Year == '2016')

A16 <-aggregate(Aridmonthdrag16$Total_ticks, by = list(Month = Aridmonthdrag16$Month), FUN = sum)
a16 <- ggplot(data = A16, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  #labs(y = 'total ticks', title = 'Arid') +
  ggtitle('2016')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) +
  annotate(geom = 'text', label = 'Arid', col = 'red',
               x = 5, y = 5.8, fontface = 'bold')

Aridmonthdrag17 <- alldrag %>%
  filter(Site == 'Arid') %>%
  group_by(Month) %>%
  filter(Year == '2017')

A17 <-aggregate(Aridmonthdrag17$Total_ticks, by = list(Month = Aridmonthdrag17$Month), FUN = sum)
a17 <-ggplot(data = A17, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2017')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Aridmonthdrag18 <- alldrag %>%
  filter(Site == 'Arid') %>%
  group_by(Month) %>%
  filter(Year == '2018')

A18 <-aggregate(Aridmonthdrag18$Total_ticks, by = list(Month = Aridmonthdrag18$Month), FUN = sum)
a18 <- ggplot(data = A18, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2018')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Aridmonthdrag19 <- alldrag %>%
  filter(Site == 'Arid') %>%
  group_by(Month) %>%
  filter(Year == '2019')

A19 <-aggregate(Aridmonthdrag19$Total_ticks, by = list(Month = Aridmonthdrag19$Month), FUN = sum)
a19 <-ggplot(data = A19,aes(x = factor(Month), y = x, group = 1))+ 
  geom_line(na.rm = TRUE)+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2019')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 
# a16
# a17
# a18
# a19

Ariddrags <- ggarrange(a16 + rremove('xlab') + rremove('ylab'), 
          a17 + rremove('xlab') + rremove('ylab'), 
          a18 + rremove('xlab') + rremove('ylab'), 
          a19 + rremove('xlab') + rremove('ylab'),
          nrow = 4, rel_heights = c(.5, .5, .5,1),
          align = 'v') 

########################### INTERMEDIATE 

Intermediatemonthdrag16 <- alldrag %>%
  filter(Site == 'Intermediate') %>%
  group_by(Month) %>%
  filter(Year == '2016')

i16 <-aggregate(Intermediatemonthdrag16$Total_ticks, by = list(Month = Intermediatemonthdrag16$Month), FUN = sum)
i16 <- ggplot(data = i16, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2016')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) + 
  annotate(geom = 'text', label = 'Intermediate', col = 'red',
               x = 5, y = 7.8, fontface = 'bold')

Intermediatemonthdrag17 <- alldrag %>%
  filter(Site == 'Intermediate') %>%
  group_by(Month) %>%
  filter(Year == '2017')

i17 <-aggregate(Intermediatemonthdrag17$Total_ticks, by = list(Month = Intermediatemonthdrag17$Month), FUN = sum)
i17 <-ggplot(data = A17, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2017')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Intermediatemonthdrag18 <- alldrag %>%
  filter(Site == 'Intermediate') %>%
  group_by(Month) %>%
  filter(Year == '2018')

i18 <-aggregate(Intermediatemonthdrag18$Total_ticks, by = list(Month = Intermediatemonthdrag18$Month), FUN = sum)
i18 <- ggplot(data = i18, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2018')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Intermediatemonthdrag19 <- alldrag %>%
  filter(Site == 'Intermediate') %>%
  group_by(Month) %>%
  filter(Year == '2019')

i19 <-aggregate(Intermediatemonthdrag19$Total_ticks, by = list(Month = Intermediatemonthdrag19$Month), FUN = sum)
i19 <-ggplot(data = i19,aes(x = factor(Month), y = x, group = 1))+ 
  geom_line(na.rm = TRUE)+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2019')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 
# a16
# a17
# a18
# a19

intermediatedrags <- ggarrange(i16 + rremove('xlab') + rremove('ylab'), 
          i17 + rremove('xlab') + rremove('ylab'), 
          i18 + rremove('xlab') + rremove('ylab'), 
          i19 + rremove('xlab') + rremove('ylab'),
          nrow = 4, rel_heights = c(.5, .5, .5,1),
          align = 'v') 





########################### MESIC

Mesicmonthdrag16 <- alldrag %>%
  filter(Site == 'Mesic') %>%
  group_by(Month) %>%
  filter(Year == '2016')

m16 <-aggregate(Mesicmonthdrag16$Total_ticks, by = list(Month = Intermediatemonthdrag16$Month), FUN = sum)
m16 <- ggplot(data = m16, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2016')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) + 
  annotate(geom = 'text', label = 'Mesic', col = 'red',
               x = 5, y = 7.8, fontface = 'bold')

Mesicmonthdrag17 <- alldrag %>%
  filter(Site == 'Mesic') %>%
  group_by(Month) %>%
  filter(Year == '2017')

m17 <-aggregate(Mesicmonthdrag17$Total_ticks, by = list(Month = Intermediatemonthdrag17$Month), FUN = sum)
m17 <-ggplot(data = m17, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2017')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Mesicmonthdrag18 <- alldrag %>%
  filter(Site == 'Mesic') %>%
  group_by(Month) %>%
  filter(Year == '2018')

m18 <-aggregate(Mesicmonthdrag18$Total_ticks, by = list(Month = Mesicmonthdrag18$Month), FUN = sum)
m18 <- ggplot(data = m18, aes(x = factor(Month), y = x, group = 1))+ 
  geom_line()+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2018')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 

Mesicmonthdrag19 <- alldrag %>%
  filter(Site == 'Mesic') %>%
  group_by(Month) %>%
  filter(Year == '2019')

m19 <-aggregate(Mesicmonthdrag19$Total_ticks, by = list(Month = Mesicmonthdrag19$Month), FUN = sum)
m19 <-ggplot(data = m19,aes(x = factor(Month), y = x, group = 1))+ 
  geom_line(na.rm = TRUE)+
  geom_point()+
  clean_background +
  xlab('Month') + ylab('Total Tick Abundance')+
  ggtitle('2019')+
  theme(axis.text = element_text(size = 6, face = 'bold'),
        axis.title = element_text( size = 2),
        plot.title = element_text(face = 'bold', hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 45))+
  scale_x_discrete(labels =c(1,2,3,4,5,6,7,8,9,
                           10,11,12)) 
# a16
# a17
# a18
# a19

mesicdrags <- ggarrange(m16 + rremove('xlab') + rremove('ylab'), 
          m17 + rremove('xlab') + rremove('ylab'), 
          m18 + rremove('xlab') + rremove('ylab'), 
          m19 + rremove('xlab') + rremove('ylab'),
          nrow = 4, rel_heights = c(.5, .5, .5,1),
          align = 'v') 


ariddrags
intermediatedrags
mesicdrags

year16 <- ggarrange(a16, i16, m16, nrow = 3)
year17 <- ggarrange(a17, i17, m17, nrow = 3)
year18 <- ggarrange(a18, i18, m18, nrow = 3)
year19 <- ggarrange(a19, i19, m19, nrow = 3)

year16
year17
year18
year19

```




Plots

twelve = total tick abundance per sampling by site and plot type
```{r tick abundance across sites with plot type fill}

twelve <-ggplot(data = aggregate,
       aes(x = Site, y = Total_ticks,
           fill = Plot))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  scale_fill_manual(values =c('darkseagreen1', 'darkseagreen3','darkseagreen4'))+
  theme_classic()+
  xlab('Site') + ylab('Total tick abundance per sampling')+
  ggtitle('Total tick abundance per sampling by site and plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) 

twelve

```

Plots

thirteen = Total tick abundance per sampling by year by plot type
```{r tick abundance by year}
aggregateyear <-aggregate(aggregateticks$Total_ticks, 
                by = list(Month = aggregateticks$Month, 
                          Site = aggregateticks$Site, 
                          Plot = aggregateticks$Plot,
                          Year = aggregateticks$Year), FUN = sum)
colnames(aggregateyear) <- c('Month', 'Site','Plot','Year', 'Total_ticks')
print(aggregateyear)

thirteen <- ggplot(data = aggregateyear,
                   aes(x = Year, y = Total_ticks,
                       fill = Plot))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  scale_fill_manual(values =c('darkseagreen1', 'darkseagreen3','darkseagreen4'))+
  clean_background +
  xlab('Year') + ylab('Total tick abundance')+
  ggtitle('Total tick abundance \nby year and plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) 

thirteen

```


Need to fix this section with some summary stats, wilcox test

######### is there a trend with total ticks across the years? ######### 
plot(alldrag$Total_ticks ~ alldrag$Year)
fit<-lm(alldrag$Total_ticks ~ alldrag$Year)
abline(fit, col = 'green', lwd = 3)

## ^^ very very small linear increase

```{r basic stats}

########## what is the distribution?  #########
hist(alldrag$Total_ticks, breaks = 5,
     xlab = 'Total ticks')
par(mfrow = c(3,1))
hist(alldrag$Total_ticks[alldrag$Site == 'Arid'],
     xlab = 'Total ticks', main = 'Arid')
hist(alldrag$Total_ticks[alldrag$Site == 'Intermediate'],
     xlab = 'Total ticks', main = 'Intermediate')
hist(alldrag$Total_ticks[alldrag$Site == 'Mesic'],
     xlab = 'Total ticks', main = 'Mesic')

# ^^^^^^ Possion distribution, not surprised


########## # can we log transform? ########### 

alldrag$log_total_ticks <- log(alldrag$Total_ticks + 1)
#view(alldrag)

hist(alldrag$log_total_ticks,
     xlab = 'Total ticks')
par(mfrow = c(3,1))
hist(alldrag$log_total_ticks[alldrag$Site == 'Arid'],
     xlab = 'Total ticks', main = 'Arid')
hist(alldrag$log_total_ticks[alldrag$Site == 'Intermediate'],
     xlab = 'Total ticks', main = 'Intermediate')
hist(alldrag$log_total_ticks[alldrag$Site == 'Mesic'],
     xlab = 'Total ticks', main = 'Mesic')

## ^^^ helped with mesic distrubution but not much else

########### how common are certain tick counts?  ############## 

## cumulative density function plot of abundance of ticks ######### 
emp_cdf <- ecdf(alldrag$Total_ticks)
plot(emp_cdf, 
     col = 'red')

## ^^ most total tick counts are less than 10, 80% are < 9 ticks total

########### is there a difference of total ticks between mesic and arid? ########### 

####### shapiro-wilk test

# H0 : data are normal
# p < 0.5 data are not normal

shapiro.test(alldrag$Total_ticks) # severly not normal, p < 2.2e-16
qqPlot(alldrag$Total_ticks) # data is not normal, do not run a t-test, like maybe CTL but no


####### lavene's test
# H0 : variance between 2 groups are equal
# if you reject lavene's test use Welch's t-test


# let's create two groups, arid or mesic

aridd <- subset(alldrag, Site == 'Arid')
mesicc <- subset(alldrag, Site == 'Mesic')

t.test(aridd$Total_ticks, mesicc$Total_ticks)

## ^ p = 0.0008068, p < 0.05, 95% does not contain 0, therefore we reject our H0 and say there is a significant difference between the groups


## paired t-test for pre-post fence (Welch Two sample t-test)
prefence <- combo %>%
  filter(year == '2016')


postfence <- combo %>%
  filter(year != '2016')

# check distributions and variance
hist(log(prefence$total_ticks + 1))
hist(log(postfence$total_ticks + 1))

# levene's test to be used for unequal variances when samples are not strictly normal
# comparing means when variances are not equal = Welch's two sample t-test
t.test(log(prefence$total_ticks + 1), log(postfence$total_ticks + 1)) # p < 2.2e-16

# wilcoxon rank sum test with continuity correction
wilcox.test(prefence$total_ticks, postfence$total_ticks) # p < 2.2e-16


```


-check to see if CTL will allow me to run ANOVAS since assumptions are not met

Tukey test is a single-step multiple comparison procedure and statistical test. It is a post-hoc analysis, which means that it is used in conjunction with an ANOVA

Tukey-Krame allows to find means of a factor that are significantly different from each other, comparing all possible pairs of means with a t-test like method


Factor(s)
  1. Site
    levels = Arid, Intermediate, Mesic
  2. Plot
    levels = Open, Partial, Total
  3. Year
    levels = 2016, 2017, 2018, 2019
  4. Month (probably less ideal way to do this)
    levels = January through December 
    
    
Plots

fouteen = ANOVA site
fifteen = ANOVA plot
sixteen = ANOVA year
seventeen = ANOVA month
```{r ANOVA for site}

# use anova() for summary 
############### ####################  SITE  ############ #######################
# H0 : all group means are equal
# HA : at least 1 mean is not equal

boxplot(Total_ticks ~ Site, 
        data = alldrag)
## ^ lots of outliers

with(alldrag, hist(Total_ticks[Site == 'Arid']))
with(alldrag, hist(Total_ticks[Site == 'Intermediate']))
with(alldrag, hist(Total_ticks[Site == 'Mesic']))

# run an ANOVA
site1 <- aov(Total_ticks ~ Site, 
        data = alldrag)
par(mfrow = c(2,2))
plot(site1)

# check residuals
res = site1$residuals
shapiro.test(res) # p << 0.05 so data not normal
hist(res)
leveneTest(Total_ticks ~ Site, 
        data = alldrag) # p << 0.05 so variances are not equal

summary(site1)
# Fvalue = 39.52
# p < 2e-16 , reject H0 and conclude that 1 site is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(site1)

# let's do this in a visual way
Post_hoc <-glht(site1, linfct = mcp(Site = 'Tukey'))
summary(Post_hoc)

############### ^ intermediate is significantly different (p < 0.001)


confint(Post_hoc) # pairwise test
cld(Post_hoc) # arid & mesic = a ; intermediate = b , so intermediate is different than arid & mesic
plot(Post_hoc) # plot 95% CIs

############ ^^^^ the mean of total tick counts is significantly different in Intermediate than Mesic and Arid!
# Intermediate has a different letter & CIs don't overlap with 0 & p-values are significant

# Site
fourteen <- ggplot(data = alldrag,
       aes(x = Site, y = Total_ticks,
           fill = Site))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('red4','red2','red'))+
  theme_classic()+
  xlab('Site') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Site')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 1.8,
           size = 9, color = 'red') +
  annotate('text', label = 'b', 
           x = 2, y = 5,
           size = 9, color = 'red') +
    annotate('text', label = 'a', 
           x = 3, y = 2.5,
           size = 9, color = 'red')


```

```{r ANOVA for plot}

anova(plot1)
######################################## PLOT  ########################################  

# H0 : all group means are equal
# HA : at least 1 mean is not equal

boxplot(Total_ticks ~ Year, 
        data = alldrag)
## ^ lots of outliers

#with(alldrag, hist(Total_ticks[Site == 'Arid']))
#with(alldrag, hist(Total_ticks[Site == 'Intermediate']))
#with(alldrag, hist(Total_ticks[Site == 'Mesic']))


plot1 <- aov(Total_ticks ~ Plot, 
        data = alldrag)
par(mfrow = c(2,2))
plot(plot1)

# check residuals
resp = plot1$residuals
shapiro.test(resp) # p << 0.05 so data not normal
hist(resp)
#leveneTest(Total_ticks ~ Year, 
        #data = alldrag) # p << 0.05 so variances are not equal


summary(plot1)
# Fvalue = 25.98 
# p < 8.33e-12  , reject H0 and conclude that 1 plot is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(plot1)

# let's do this in a visual way
Post_hocp <-glht(plot1, linfct = mcp(Plot = 'Tukey'))
summary(Post_hocp)

confint(Post_hocp) # pairwise test
cld(Post_hocp) # different letters are different from each other open is different than partial & total!
plot(Post_hocp) # plot 95% CIs

########################  ^^^^ Open is significantly different

# Open has a different letter & CIs don't overlap with 0 & p-values are significant

# Plot
fifteen <-ggplot(data = alldrag,
       aes(x = Plot, y = Total_ticks,
           fill = Plot))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme_classic()+
  xlab('Plot') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Plot')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 1.6,
           size = 9, color = 'red') +
  annotate('text', label = 'b', 
           x = 2, y = 3.4,
           size = 9, color = 'red') +
    annotate('text', label = 'b', 
           x = 3, y = 4.3,
           size = 9, color = 'red')


```

```{r ANOVA for year}

anova(year1)
################################################## YEAR  ######################################## 

# H0 : all group means are equal
# HA : at least 1 mean is not equal

boxplot(Total_ticks ~ Year, 
        data = alldrag)
## ^ lots of outliers

#with(alldrag, hist(Total_ticks[Site == 'Arid']))
#with(alldrag, hist(Total_ticks[Site == 'Intermediate']))
#with(alldrag, hist(Total_ticks[Site == 'Mesic']))


year1 <- aov(Total_ticks ~ Year, 
        data = alldrag)
par(mfrow = c(2,2))
plot(year1)

# check residuals
resy = year1$residuals
shapiro.test(resy) # p << 0.05 so data not normal
hist(resy)
#leveneTest(Total_ticks ~ Year, 
        #data = alldrag) # p << 0.05 so variances are not equal

summary(year1)
# Fvalue = 21.63
# p  = 8.33e-12, p < 0.001 reject H0 and conclude that 1 site is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)

# need to transform year into factor to run TukeyHSD
class(alldrag$Year) # integer
alldrag$Year = as.factor(alldrag$Year)
class(alldrag$Year) #factor

TukeyHSD(year1)

# let's do this in a visual way
Post_hocy <-glht(year1, linfct = mcp(Year = 'Tukey'))
summary(Post_hocy)

confint(Post_hocy) # pairwise test
cld(Post_hocy) # different letters are different from each other
plot(Post_hocy) # plot 95% CIs

#### ^^^^ 2017 & 2018 similar 'c', 2016 different 'a', 2019 different 'b'

# 2016, 2019, 2017 & 2018 has a different letter & CIs don't overlap with 0 & p-values are significant


# Plot
sixteen<- ggplot(data = alldrag,
       aes(x = Year, y = Total_ticks,
           fill = Year))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('lightblue1','lightblue2','lightblue3','lightblue4'))+
  theme_classic()+
  xlab('Year') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Year')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = .5,
           size = 9, color = 'red') +
  annotate('text', label = 'c', 
           x = 2, y = 3.93,
           size = 9, color = 'red') +
    annotate('text', label = 'c', 
           x = 3, y = 3.9,
           size = 9, color = 'red')+
   annotate('text', label = 'b', 
           x = 4, y = 2.72,
           size = 9, color = 'red')




```

```{r ANOVA for month}

anova(month1)
################### ########## # MONTH  ############# ########## #######

# H0 : all group means are equal
# HA : at least 1 mean is not equal

boxplot(Total_ticks ~ Month, 
        data = alldrag)
## ^ lots of outliers

#with(alldrag, hist(Total_ticks[Site == 'Arid']))
#with(alldrag, hist(Total_ticks[Site == 'Intermediate']))
#with(alldrag, hist(Total_ticks[Site == 'Mesic']))


month1 <- aov(Total_ticks ~ Month, 
        data = alldrag)
par(mfrow = c(2,2))
plot(month1)

# check residuals
resm = month1$residuals
shapiro.test(resm) # p << 0.05 so data not normal
hist(resm)
leveneTest(Total_ticks ~ Month, 
        data = alldrag) # p << 0.05 so variances are not equal

summary(month1)
# Fvalue = 24.85
# p < 2e-16 , reject H0 and conclude that 1 month is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(month1)

par(mfrow = c(1,1))
# let's do this in a visual way
Post_hocm <-glht(month1, linfct = mcp(Month = 'Tukey'))
summary(Post_hocm)

confint(Post_hocm) # pairwise test
cld(Post_hocm) # different letters are different from each other
plot(Post_hocm) # plot 95% CIs

#### ^^^^ Months (May?) is significantly different, this is kinda  poor way of analyzing months, maybe group by season but that doesn't say much either

seventeen <- ggplot(data = alldrag,
       aes(x = Month, y = Total_ticks))+
  geom_bar(stat = 'summary', fun.y = 'mean', fill = 'grey')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  theme_classic()+
  xlab('Month') + ylab('Total Ticks')+
  ggtitle('Total Ticks by Month')+
  theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = .5,
           size = 7, color = 'red') +
  annotate('text', label = 'ab', 
           x = 2, y = .7,
           size = 7, color = 'red') +
    annotate('text', label = 'ac', 
           x = 3, y = 3.2,
           size = 7, color = 'red')+
   annotate('text', label = 'e', 
           x = 4, y = 6.6,
           size = 7, color = 'red') +
  annotate('text', label = 'de', 
           x = 5, y = 6.4,
           size = 7, color = 'red') +
  annotate('text', label = 'cd', 
           x = 6, y = 4.7,
           size = 7, color = 'red') +
    annotate('text', label = 'ab', 
           x = 7, y = 2.1,
           size = 7, color = 'red')+
   annotate('text', label = 'a', 
           x = 8, y = .5,
           size = 7, color = 'red')+
   annotate('text', label = 'a', 
           x = 9, y = .5,
           size = 7, color = 'red')+ 
  annotate('text', label = 'bc', 
           x = 10, y = 3.6,
           size = 7, color = 'red')+
  annotate('text', label = 'ac', 
           x = 11, y = 3.2,
           size = 7, color = 'red')+
  annotate('text', label = 'a', 
           x = 12, y = .5,
           size = 7, color = 'red')


```

```{r plot of anovas}
library(effects)
plot(predictorEffects(plot1, ~ Plot),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))

plot(predictorEffects(site1, ~ Site),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))

plot(predictorEffects(year1, ~ Year),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
# 
# ggplot(data = combo, aes(x = site, y = total_ticks,
#         fill = plot, color = plot)) + 
#     geom_jitter(size = 3, position = position_dodge2(width = 0.3)) +
#     labs(x = "YEAR", y = "COUNT") +
#     theme_classic()+
#     theme(text = element_text(size = 14)) 
```



##What are good predictors (X) of total tick abundance (Y)?

Ran *generalized linear models* because data breaks assumptions of normality. Used family poisson because of count data.

Object *'tickmodels'* is table of df, AIC, BIC, variables used in glm models
- lowest AIC: all 4 variables (Site, Year, Plot, Month)
- lowest AIC within single predictors = Year
  Year AIC = 10496
  Site AIC = 10665
  Plot AIC = 10871

tick1 - tick7 glm models

Graph
  eighteen = kable table with df, AIC, BIC, variables
  nineteen = table with p values

```{r glm for tick data}

tick1 <- glm(Total_ticks ~ Year, 
               data = alldrag,
             family = 'poisson')

tick2 <- glm(Total_ticks ~ Site, 
               data = alldrag,
              family = 'poisson')

tick3 <- glm(Total_ticks ~ Plot, 
               data = alldrag,
              family = 'poisson')

tick4 <- glm(Total_ticks ~ Month, 
               data = alldrag,
              family = 'poisson')

tick5 <- glm(Total_ticks ~ Site + Year, 
               data = alldrag,
              family = 'poisson')

tick6 <- glm(Total_ticks ~ Site + Year + Plot, 
               data = alldrag,
              family = 'poisson')

tick7 <- glm(Total_ticks ~ Site + Year + Plot + Month, 
               data = alldrag,
              family = 'poisson')


# calculate AIC of each model

tickresults <- AIC(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7)

# add other metrics to table
tickmodels <- list(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
tickresults$BIC <- sapply(tickmodels, BIC) # add column for BIC

tick_model_summary <- lapply(tickmodels, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(tickmodels)) {
  tickresults$rsq[i] <- tick_model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  tickresults$adj_rsq[i] <- tick_model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
tickresults$variables <- c('Year','Site', 'Plot', 'Month', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

# make a nice table!
tickmodels <-kable(tickresults, digits = 2, align = 'c') %>%
  kable_styling()

eighteen <-tickmodels


nineteen <- export_summs(tick1, tick2, tick3, scale = TRUE)
#min(tickresults$AIC)

```

Kruskal-Wallis test
- non-parametric test similar to a single factor anova
- uses the ranks of data points
```{r kruskal-wallis test}

kw1 <- kruskal.test(total_ticks ~ site, data = combo) # p = 1.165e-09
kw2 <- kruskal.test(total_ticks ~ plot, data = combo) # p = 2.601e-12
kw3 <-kruskal.test(total_ticks ~ year, data = combo) # p < 2.2e-16

library(rstatix)
plot.kw(wallisresults, strip = TRUE, posthoc = TRUE, adjust = 'bonferroni')

Kw1 <- combo %>% add_xy_position(x = 'site')
ggboxplot(combo, x = 'site', y = 'total_ticks')+
  stat_pvalue_manual(combo, hide.ns = TRUE) +
  labs(subtitle = get_test_label(res.kruksal, detailed = TRUE),
       caption = get_pwc_label(pwc))




```

Ran *generalized linear mixed-effect models* because data breaks assumptions of normality. Used family poisson because of count data. *Incorporating mixed-effects.*

Object *'tick_glmm_models'* is table of df, AIC, BIC, variables used in glmm models
- lowest AIC: all 4 variables (Site, Year, Plot, Month) *[same as glm models]*
- lowest AIC within single predictors = Year *[same as glm models]*
  Year AIC = 9958 *[lower AIC than glm models]*
  Site AIC = 10127 *[lower AIC than glm models]*
  Plot AIC = 10333 *[lower AIC than glm models]*

tick8 - tick14 glm models

Graph
  twenty = kable table with df, AIC, BIC, variables
  twentyone = table with p values

```{r glmm tick models}

tick8 <- glmer(Total_ticks ~ Year + (1|Block), 
               data = alldrag,
             family = poisson(link = 'log'))

tick9 <- glmer(Total_ticks ~ Site + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))

tick10 <- glmer(Total_ticks ~ Plot + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))

tick11 <- glmer(Total_ticks ~ Month + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))

tick12 <- glmer(Total_ticks ~ Site + Year + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))

tick13 <- glmer(Total_ticks ~ Site + Year + Plot + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))

tick14 <- glmer(Total_ticks ~ Site + Year + Plot + Month + (1|Block), 
               data = alldrag,
              family = poisson(link = 'log'))


# calculate AIC of each model

tick_glmm_results <- AIC(tick8, tick9, tick10,
                    tick11, tick12, tick13,
                   tick14)

# add other metrics to table
tick_glmm_models <- list(tick8, tick9, tick10,
                    tick11, tick12, tick13,
                   tick14) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
tick_glmm_results$BIC <- sapply(tick_glmm_models, BIC) # add column for BIC

tick_glmm_model_summary <- lapply(tick_glmm_models, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(tick_glmm_models)) {
  tick_glmm_results$rsq[i] <- tick_glmm_model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  tick_glmm_results$adj_rsq[i] <- tick_glmm_model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
tick_glmm_results$variables <- c('Year','Site', 'Plot', 'Month', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

# add another column with random effects
tick_glmm_results$randomeffect <- c('Block','Block','Block','Block','Block','Block','Block')

# make a nice table!
tick_glmm_models <-kable(tick_glmm_results, digits = 2, align = 'c') %>%
  kable_styling()

twenty <- tick_glmm_models

twentyone <- export_summs(tick8,tick9,tick10,tick11, tick12, tick13, tick14, scale = TRUE)

```

I need to check if I can run linear regression (lm) models since CTL would apply to this dataframe.

Models
  tick15
    Total_ticks ~ Plot
    Plot is a significant predictor of total ticks 
        - p = 8.89e-06
        - adjusted r^2 = 0.0346 ; only 3.59% variation in the total number of ticks can be explained by plot type 
        
  tick16
    Total_ticks ~ Site
    Site is a significant predictor of total ticks 
        - p = 1.08e-06
        - adjusted r^2 = 0.05237 ; only 5.23% variation in the total number of ticks can be explained by plot type 
        
Graph
  twentytwo = table with p values

```{r linear regression of ticks}
## don't have two numeric variables to compare to so can't do correlation on this dataset, maybe later when I have temp data

# check for linearity, I know I shouldn't have categorical variables in there
plot.drag <-data.frame(alldrag[,2:3],alldrag[,5:6])
pairs.panels(plot.drag, density = TRUE, cor = FALSE, lm = TRUE)

#################### run linear regression models ##########################

####### Total_ticks ~ PLOT #######
tick15 <- lm(Total_ticks ~ Plot, data = alldrag)
summary(tick15)
# intercept (b0) : 1.1247  ; pvalue = 8.89e-06 [Plot Open]
# plot partial (b1) : 1.7097 ; pvalue = 1.82e-06
# plot total (b2) : 2.5183 ; pvalue = 2.62e-12
### ^^ for every one unit increase in plot Total, there is a corresponding 2.5183 increase in number of ticks
### my linear model is Yi = 1.1247 + 2.5183Xi

# r^2 = Multiple R-squared = 0.0359	; Adjusted R-squared = 0.0346
## ^^ 3.59% of the variation in the total number of ticks (Y) can be explained by the plot type (X)


####### Total_ticks ~ SITE #######
tick16 <- lm(Total_ticks ~ Site, data = alldrag)
summary(tick16)
# intercept (b0) : 1.2444   ; pvalue = 1.08e-06 [Site Arid]
# site intermediate (b1) : 3.0155 ; pvalue = < 2e-16
# site mesic (b2) : 0.7705 ; pvalue = 0.0305
### ^^ for every one unit increase in site Arid, there is a corresponding 2.5183 increase in number of ticks
### my linear model is Yi = 1.2444 + 3.0155Xi

# r^2 = Multiple R-squared =  0.05373	; Adjusted R-squared = 0.05237 
## ^^ 5.37% of the variation in the total number of ticks (Y) can be explained by the site type (X)

# put these models in nice graph
summ(tick15, confint = TRUE)
summ(tick16, confint = TRUE)

# export both graphs
twentytwo <- export_summs(tick15, tick16, scale = TRUE)


```



###Mammal Data


Excel sheets called 
- 'tejon_mammals1819_09022020'
- 'tejon_mammalcount_09032020'


in these datasheets I name columns with lowercase compared to tick data sheets (i.e. month not Month)
```{r import data frames and organize}

# full dataset
mammals <- read.csv("~/Desktop/tejon_mammals1819_09022020.csv")

# subset for important
mammal <-data.frame(mammals[,4:6],mammals[,9])

# numeric dataframe for community analysis
Count <- read.csv("~/Desktop/tejon_mammalcount_09032020.csv", header = TRUE)

# remove the arid with NAs, this will be the full dataset with species richenss and diversity added
count <- Count[-4,]

# keep only numeric data to run community analyses
numeric <- count[,c(-1,-2,-9,-10)]

# organize data into right forms
count$plot <- factor(count$plot, 
                        levels = c('open','partial','total'))


# species richness
sppr <- specnumber(numeric)

# shannon diversity
shannondiv <- diversity(numeric)

# add species richness to count
count$sppr<-specnumber(numeric)

# add shannon diversity to count
count$shannondiv <- diversity(numeric)

pal <- c("lightsalmon1", "gold1", "palegreen4")
plotcolor <- pal




```

Graphs


Total animal

twentythree = total animal abundance by site (Tukey Kramer values)
twentyfour = total animal abundance by plot
twentyfive = total animal abundance by site with plot fill
twentysix = total animal abundance by site and plot combo (I like this one)
```{r}
########### TOTAL ANIMAL (Y) ############

# total animal abundance by site
twentythree <-ggplot(data = count,
       aes(x = site, y = total,
           fill = site))+
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('red4','red2','red'))+
  clean_background +
  xlab('Site Types') + ylab('Animal Abundance')+
  ggtitle('Animal Abundance by Site')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none") +
  annotate('text', label = 'a', 
           x = 1, y = 42,
           size = 9, color = 'red') +
  annotate('text', label = 'a', 
           x = 2, y = 23,
           size = 9, color = 'red') +
    annotate('text', label = 'b', 
           x = 3, y = 72,
           size = 9, color = 'red')

# total animal abundance by plot
twentyfour <-ggplot(data = count,
       aes(x = plot, y = total,
       fill = plot))+
  scale_x_discrete(labels = c("Open", "Partial", "Total")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  clean_background +
  xlab('Plot Types') + ylab('Animal Abundance')+
  ggtitle('Animal Abundance by Plot')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none")  



##  total animal abundance by site with fill
twentyfive <-ggplot(data = count,
       aes(x = site, y = total,
           fill = plot))+
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  clean_background +
  xlab('Site Types') + ylab('Animal Abundance')+
  ggtitle('Animal Abundance by Site')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5)) 



twentysix <- ggarrange(twentythree + rremove('legend'), twentyfour + rremove('ylab') +rremove('legend'),ncol = 2)

twentythree
twentyfour
twentyfive
twentysix
```

Graphs

twentyseven = species richness by site type (boxplot)
twentyeight = species richness by plot type (boxplot)
twentynine = species richness by site type with error bars
thirty = species richness by plot type with error bars

```{r species richness by site and plot}

############################## SITE ##############################
## species richness by site
twentyseven <- ggplot(count, aes(x = site, y = sppr))+
  geom_boxplot(fill = sitecolor) +
  coord_cartesian(ylim = c(0,6)) +
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  clean_background + 
  labs(x = "Site Types",
       y = "Species Richness",
       title = "Species Richness by Site") +
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5))


############################## PLOT ##############################

## species richness by plot

twentyeight <- ggplot(count, aes(x = plot, y = sppr))+
  geom_boxplot(fill = plotcolor) +
  coord_cartesian(ylim = c(0,6)) +
  scale_x_discrete(labels = c("Open", "Partial", "Total")) +
  clean_background + 
  labs(x = "Site Types",
       y = "Species Richness",
       title = "Species Richness by Plot") +
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5))

### geom_bar with error bars

# species diversity by site with error bars

twentynine <- ggplot(data = count,
       aes(x = site, y = sppr,
           fill = site))+
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('red4','red2','red'))+
  clean_background +
  xlab('Site Types') + ylab('Spieces Richness')+
  ggtitle('Species Richness by Site')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none") 



# species diversity by plot with error bars

thirty <- ggplot(data = count,
       aes(x = plot, y = sppr,
           fill = plot))+
  scale_x_discrete(labels = c("Open", "Partial", "Total")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  clean_background +
  xlab('Plot Types') + ylab('Spieces Richness')+
  ggtitle('Species Richness by Plot')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none") 

twentyseven
twentyeight
twentynine
thirty
```

Shannon Diversity

Plots

thirtyone = shannon diversity by site (boxplot)
thirtytwo = shannon diversity by plot (boxplot)
thirtythree = shannon diversity by site (Tukey Kramer values)
thirtyfour = shannon diversity by plot
```{r}

## shannon diversity by site
thirtyone <- ggplot(count, aes(x = site, y = shannondiv))+
  geom_boxplot(fill = sitecolor) +
  coord_cartesian(ylim = c(0,1.5)) +
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  clean_background + 
  labs(x = "Site Types",
       y = "Shannon Diversity",
       title = "Shannon Diversity by Site") +
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5))

## shannon diversity by plot

thirtytwo <- ggplot(count, aes(x = plot, y = shannondiv))+
  geom_boxplot(fill = plotcolor) +
  coord_cartesian(ylim = c(0,1.5)) +
  scale_x_discrete(labels = c("Open", "Partial", "Total")) +
  clean_background + 
  labs(x = "Site Types",
       y = "Shannon Diversity",
       title = "Shannon Diversity by Plot") +
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5))

### geom_bar with error bars

# shannon diversity by site with error bars

thirtythree <- ggplot(data = count,
       aes(x = site, y = shannondiv,
           fill = site))+
  scale_x_discrete(labels = c("Arid", "Intermediate", "Mesic")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('red4','red2','red'))+
  clean_background +
  xlab('Site Types') + ylab('Shannon Diversity')+
  ggtitle('Shannon Diversity by Site')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none") +
  annotate('text', label = 'ab', 
           x = 1, y = 1.1,
           size = 9, color = 'red') +
  annotate('text', label = 'a', 
           x = 2, y = .7,
           size = 9, color = 'red') +
    annotate('text', label = 'b', 
           x = 3, y = 1.3,
           size = 9, color = 'red') 



# shannon diversity by plot with error bars

thirtyfour <- ggplot(data = count,
       aes(x = plot, y = shannondiv,
           fill = plot))+
  scale_x_discrete(labels = c("Open", "Partial", "Total")) +
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  clean_background +
  xlab('Plot Types') + ylab('Shannon Diversity')+
  ggtitle('Shannon Diversity by Plot')+
  theme(axis.text = element_text(size = 12, face = 'bold'),
        axis.title = element_text( size = 14),
        plot.title = element_text(face = 'bold', hjust = 0.5),
        legend.position = "none") 

thirtyone
thirtytwo
thirtythree
thirtyfour
```

ANOVA 

1. plot type (X)
    mammal 1 : abundance (Y) 
    mammal 2 : shannon diversity (Y) 
    mammal 3 : species richness (Y) 
2. site type 
    mammal 4 : abundance (Y) 
    mammal 5 : shannon diversity (Y) 
    mammal 6 : species richness (Y) 
```{r ANOVA}

# real quick

### plot on mammal metrics
mammal1 <- aov(total ~ plot, data = count)
summary(mammal1) # p value = 0.597

mammal2 <- aov(sppr ~ plot, data = count)
summary(mammal2) # p value = 0.422

mammal3 <-aov(shannondiv ~ plot, data = count)
summary(mammal3) # p value = 0.501

### site on mammal metrics
mammal4 <- aov(total ~ site, data = count)
summary(mammal4) # p value = 0.00857 [SIGNIFICANT]

mammal5 <- aov(sppr ~ site, data = count)
summary(mammal5) # p value = 0.0723

mammal6 <-aov(shannondiv ~ site, data = count)
summary(mammal6) # p value = 0.0485 [SIGNIFICANT]



## mammal4 and mammal6 are signficant so I will check residuals and run a Tukey Kramer


########## Mammal4 (total ~ site)
# run an ANOVA
plot(mammal4)

# check residuals
res4 = mammal4$residuals
shapiro.test(res4) # p = 0.617 (p > 0.05) so data is normal
hist(res4)
leveneTest(total ~ site, data = count) # p = 0.781, (p > 0.05) so variances are equal

summary(mammal4)
# Fvalue = 11.66
# p = 0.00857 (p < 0.05), reject H0 and conclude that 1 site is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(mammal4)

# let's do this in a visual way
Post_hoc4 <-glht(mammal4, linfct = mcp(site = 'Tukey'))
summary(Post_hoc4)

############### ^ mesic is significantly different (p < 0.001)


confint(Post_hoc4) # pairwise test
cld(Post_hoc4) # arid & intermediate = a ; mesic = b , so intermediate is different than arid & mesic
plot(Post_hoc4) # plot 95% CIs

############ ^^^^ the total mammal counts is significantly different in Mesic than Intermediate and Arid!
# Mesic has a different letter & CIs don't overlap with 0 & p-values are significant




########## Mammal6 (shannondiv ~ site)
# run an ANOVA
plot(mammal6)

# check residuals
res6 = mammal6$residuals
shapiro.test(res6) # p = 0.4144 (p > 0.05) so data is normal
hist(res6)
leveneTest(shannondiv ~ site, data = count) # p = 0.3411, (p > 0.05) so variances are equal

summary(mammal6)
# Fvalue = 5.226
# p = 0.0485 (p < 0.05), reject H0 and conclude that 1 site is significantly different

## ANOVA told us 1 mean is significantly different from the other
# Run a Tukey Kramer to avoid Type 1 error (running a bunch of t-tests)

# Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(mammal6)

# let's do this in a visual way
Post_hoc6 <-glht(mammal6, linfct = mcp(site = 'Tukey'))
summary(Post_hoc6)



confint(Post_hoc6) # pairwise test
cld(Post_hoc6) # arid = ab, & intermediate = a ; mesic = b , so intermediate and mesic are different 
plot(Post_hoc6) # plot 95% CIs

############ ^^^^ the shannon diversity is significantly different between intermediate and mesic but these results are not as clear 
# intermediate has a different letter & CIs don't overlap with 0 & p-values are significant



```

PCA plot
- honestly these are not as clear, but willing to keep working on them if people are interested

Graphs

PCA1 = count by species
  site 1 - 3 : arid
  site 4 - 7 : intermediate
  site 8 - 10 : mesic
```{r PCA plots - needs more work}

################### Attempt two ###################

mam.pca <- prcomp(numeric, center = TRUE, scale = TRUE)
print(mam.pca)
plot(mam.pca, type = 'l')
summary(mam.pca)

out <- as.data.frame(mam.pca$x) #get the PC values for each observation so we can plot it

out$site <-c('arid','arid','arid',
                 'intermediate','intermediate','intermediate',
                 'mesic','mesic','mesic')
PCA1 <- ggplot(out, aes(x=PC1, y=PC2, color = site))+
  clean_background+
  geom_point()

PCA1

```

################ COMBO DATASETS ################


Excel sheet 'tejon_tickmammal_09032020'

Has tick and mammmal data combined with calculated species richness and shannondiversity metrics associated with each site
```{r read in datasets and organize}
combo <- read.csv("~/Desktop/tejon_tickmammal_09032020.csv")

combo$site <- factor(combo$site, 
                        levels = c('Arid','Intermediate','Mesic'))

combo$plot <- factor(combo$plot, 
                        levels = c('Open','Partial','Total'))
combo$month <- factor(combo$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
combo$year = as.factor(combo$year)


```

**Linear Regression models - may not be appropriate for these data**

models : tick14 - tick19
- Lowest AIC for *multivariables* is tick18 (total_tick ~ site, year, plot) 
- Lowest AIC for *singularvariable* is tick14 (total_tick ~ year)

models : mammal7 - mammal12
- Lowest AIC for *multivariables* is mammal11 (total_mammal ~ site, year, plot) 
- Lowest AIC for *singularvariable* is mammal (total_mammal ~ site)


models : combo1 - combo6
- Lowest AIC for *multivariables* is combo5 (total_tick ~ site, year, plot, shannon) 
- Lowest AIC for *singularvariable* is combo2 (total_tick ~ mammal species richness) [not very different between tot, richness, shannon AICs]


Graphs/ Tables with df, AIC, BIC, rsq, adj_rsq, variables
1. tickmodel_lm 
2. mammalmodel_lm 
3. combomodel_lm
```{r linear regression models}
####################### TICKS ###########################
##### run individual models on total mammal abundance
tick14 <- lm(total_ticks ~ year, 
               data = combo)

tick15 <- lm(total_mammal ~ site, 
               data = combo)

tick16 <- lm(total_mammal ~ plot, 
               data = combo)

tick17 <- lm(total_mammal ~ site + year, 
               data = combo)

tick18  <- lm(total_mammal ~ site + year + plot, 
               data = combo)

tick19 <- lm(total_mammal ~ site + year + plot + month, 
               data = combo)


# calculate AIC of each model

result_lm_tick <- AIC(tick14, tick15, tick16,
                        tick17, tick18, tick19)

# add other metrics to table
model_lm_tick <- list(tick14, tick15, tick16,
                        tick17, tick18, tick19) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_lm_tick$BIC <- sapply(model_lm_tick, BIC) # add column for BIC

model_lm_tick_summary <- lapply(model_lm_tick, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_lm_tick)) {
  result_lm_tick$rsq[i] <- model_lm_tick_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_lm_tick$adj_rsq[i] <- model_lm_tick_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_lm_tick$variables <- c('Year','Site', 'Plot', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

tickmodel_lm <-kable(result_lm_tick, digits = 2, align = 'c',
                        caption = '<b>Linear regression models for total tick abundance</b>', 
                        format = 'html') %>%
  kable_styling()

tickmodel_lm 




####################### MAMMALS ###########################
##### run individual models on total mammal abundance
mammal7 <- lm(total_mammal ~ year, 
               data = combo)

mammal8 <- lm(total_mammal ~ site, 
               data = combo)

mammal9 <- lm(total_mammal ~ plot, 
               data = combo)

mammal10 <- lm(total_mammal ~ site + year, 
               data = combo)

mammal11 <- lm(total_mammal ~ site + year + plot, 
               data = combo)

mammal12 <- lm(total_mammal ~ site + year + plot + month, 
               data = combo)


# calculate AIC of each model

result_lm_mammal <- AIC(mammal7, mammal8,mammal9,
               mammal10, mammal11, mammal12)

# add other metrics to table
model_lm_mammal <- list(mammal7, mammal8,mammal9,
               mammal10, mammal11, mammal12) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_lm_mammal$BIC <- sapply(model_lm_mammal, BIC) # add column for BIC

model_lm_mammal_summary <- lapply(model_lm_mammal, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_lm_mammal)) {
  result_lm_mammal$rsq[i] <- model_lm_mammal_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_lm_mammal$adj_rsq[i] <- model_lm_mammal_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_lm_mammal$variables <- c('Year','Site', 'Plot', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

mammalmodel_lm <-kable(result_lm_mammal, digits = 2, align = 'c',
                        caption = '<b>Linear regression models for total mammal abundance</b>', 
                        format = 'html') %>%
  kable_styling()

mammalmodel_lm 


####################### COMBINATION OF TICKS AND MAMMALS ###########################

combo1 <- lm(total_ticks ~ total_mammal, 
               data = combo)

combo2 <- lm(total_ticks ~ rich_mammal, 
               data = combo)

combo3 <- lm(total_ticks ~ shan_mammal, 
               data = combo)

combo4 <- lm(total_ticks ~ shan_mammal + total_mammal, 
               data = combo)

combo5 <- lm(total_ticks ~ site + year + plot + month + shan_mammal, 
               data = combo)

combo6 <- lm(total_ticks ~ site + year + plot + month + shan_mammal + total_mammal, 
               data = combo)


# calculate AIC of each model

result_lm_combo <- AIC(combo1,combo2, combo3,combo4, combo5,
                        combo6)

# add other metrics to table
model_lm_combo <- list(combo1,combo2, combo3,combo4, combo5,
                        combo6) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_lm_combo$BIC <- sapply(model_lm_combo, BIC) # add column for BIC

model_lm_combo_summary <- lapply(model_lm_combo, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_lm_combo)) {
  result_lm_combo$rsq[i] <- model_lm_combo_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_lm_combo$adj_rsq[i] <- model_lm_combo_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_lm_combo$variables <- c('Total Mammal','Mammal Species Richness', 'Mammal Shannon Diversity', 'Shannon + Total', 'Site + Year + Plot + Shannon', 'Site + Year + Plot + Shannon + Total')

combomodel_lm <-kable(result_lm_combo, digits = 2, align = 'c',
                        caption = '<b>Linear regression models for tick abundance with mammal variables</b>', 
                        format = 'html') %>%
  kable_styling()

combomodel_lm 

```

**Generalized Linear Regression models**

models : tick20 - tick25
- Lowest AIC for *multivariables* is tick24 (total_tick ~ site, year, plot) [same as lm]
- Lowest AIC for *singularvariable* is tick20 (total_tick ~ year) [very clear better singular predictor]

models : mammal13 - mammal18
- Lowest AIC for *multivariables* is mammal17 (total_mammal ~ site, year, plot) [same as lm]
- Lowest AIC for *singularvariable* is mammal14 (total_mammal ~ site) [very clear better singular predictor]


models : combo7 - combo12
- Lowest AIC for *multivariables* is combo12 (total_tick ~ site, year, plot, shannon, total) [different than lm] 
- Lowest AIC for *singularvariable* is combo28 (total_tick ~ mammal species richness) [different than lm]


Graphs/ Tables with df, AIC, BIC, rsq, adj_rsq, variables
1. tickmodel_glm 
2. mammalmodel_glm 
3. combomodel_glm

```{r}
####################### TICKS ###########################
##### run individual models on total mammal abundance

library(lme4)
tick20 <- glm(total_ticks ~ year, 
               data = combo,
              family = 'poisson')

tick21 <- glm(total_mammal ~ site, 
               data = combo,
              family = 'poisson')

tick22 <- glm(total_mammal ~ plot, 
               data = combo,
              family = 'poisson')

tick23 <- glm(total_mammal ~ site + year, 
               data = combo,
              family = 'poisson')

tick24  <- glm(total_mammal ~ site + year + plot, 
               data = combo,
              family = 'poisson')

tick25 <- glm(total_mammal ~ site + year + plot + month, 
               data = combo,
              family = 'poisson')


# calculate AIC of each model

result_glm_tick <- AIC(tick20, tick21, tick22,
                        tick23, tick24, tick25)

# add other metrics to table
model_glm_tick <- list(tick20, tick21, tick22,
                        tick23, tick24, tick25) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_glm_tick$BIC <- sapply(model_glm_tick, BIC) # add column for BIC

model_glm_tick_summary <- lapply(model_glm_tick, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_glm_tick)) {
  result_glm_tick$rsq[i] <- model_glm_tick_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_glm_tick$adj_rsq[i] <- model_glm_tick_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_glm_tick$variables <- c('Year','Site', 'Plot', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

tickmodel_glm <-kable(result_glm_tick, digits = 2, align = 'c',
                        caption = '<b>Generalized linear regression models for total tick abundance</b>', 
                        format = 'html') %>%
  kable_styling()

tickmodel_glm 




####################### MAMMALS ###########################
##### run individual models on total mammal abundance
mammal13 <- glm(total_mammal ~ year, 
               data = combo,
              family = 'poisson')

mammal14 <- glm(total_mammal ~ site, 
               data = combo,
              family = 'poisson')

mammal15 <- glm(total_mammal ~ plot, 
               data = combo,
              family = 'poisson')

mammal16 <- glm(total_mammal ~ site + year, 
               data = combo,
              family = 'poisson')

mammal17 <- glm(total_mammal ~ site + year + plot, 
               data = combo,
              family = 'poisson')

mammal18 <- glm(total_mammal ~ site + year + plot + month, 
               data = combo,
              family = 'poisson')


# calculate AIC of each model

result_glm_mammal <- AIC(mammal13, mammal14,mammal15,
               mammal16, mammal17, mammal18)

# add other metrics to table
model_glm_mammal <- list(mammal13, mammal14,mammal15,
               mammal16, mammal17, mammal18) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_glm_mammal$BIC <- sapply(model_glm_mammal, BIC) # add column for BIC

model_glm_mammal_summary <- lapply(model_glm_mammal, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_glm_mammal)) {
  result_glm_mammal$rsq[i] <- model_glm_mammal_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_glm_mammal$adj_rsq[i] <- model_glm_mammal_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_glm_mammal$variables <- c('Year','Site', 'Plot', 'Site + Year', 'Site + Year + Plot', 'Site + Year + Plot + Month')

mammalmodel_glm <-kable(result_glm_mammal, digits = 2, align = 'c',
                        caption = '<b>Generalized linear regression models for total mammal abundance</b>', 
                        format = 'html') %>%
  kable_styling()

mammalmodel_glm 


####################### COMBINATION OF TICKS AND MAMMALS ###########################

combo7 <- glm(total_ticks ~ total_mammal, 
               data = combo,
              family = 'poisson')

combo8 <- glm(total_ticks ~ rich_mammal, 
               data = combo,
              family = 'poisson')

combo9 <- glm(total_ticks ~ shan_mammal, 
               data = combo,
              family = 'poisson')

combo10 <- glm(total_ticks ~ shan_mammal + total_mammal, 
               data = combo,
              family = 'poisson')

combo11 <- glm(total_ticks ~ site + year + plot + month + shan_mammal, 
               data = combo,
              family = 'poisson')

combo12 <- glm(total_ticks ~ site + year + plot + month + shan_mammal + total_mammal, 
               data = combo,
              family = 'poisson')


# calculate AIC of each model

result_glm_combo <- AIC(combo7,combo8, combo9,combo10, combo11,
                        combo12)

# add other metrics to table
model_glm_combo <- list(combo7,combo8, combo9,combo10, combo11,
                        combo12) # keep models in the same order as they were created

# BIC places higher penalty on extra parameter
result_glm_combo$BIC <- sapply(model_glm_combo, BIC) # add column for BIC

model_glm_combo_summary <- lapply(model_glm_combo, summary)


# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(model_glm_combo)) {
  result_glm_combo$rsq[i] <- model_glm_combo_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result_glm_combo$adj_rsq[i] <- model_glm_combo_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again

# add column with variables used in model
result_glm_combo$variables <- c('Total Mammal','Mammal Species Richness', 'Mammal Shannon Diversity', 'Shannon + Total', 'Site + Year + Plot + Shannon', 'Site + Year + Plot + Shannon + Total')

combomodel_glm <-kable(result_glm_combo, digits = 2, align = 'c',
                        caption = '<b>Generalized linear regression models for tick abundance with mammal variables</b>', 
                        format = 'html') %>%
  kable_styling()

combomodel_glm 



```

## Zero Inflated hurdle models

Neither total tick or mammal counts are normally distributed (shapiro.test p << 0.05, meaning data not normal).

The models are based on Poisson and Negative binomial distribution are expecting count data (integer values only).

Hurdle model: treat as 2 part process
  a. 1st part is 0
  b. 2nd part is non-zero
  
Zero inflated Poisson (ZIP): 
  Two types of 0s : good and bad
    variance = mean

Zero inflated negative binomial (ZINB):

  
Negative binomial version is more significant than poisson

Models
*1. ZIP_round_down*
  Site was a significant predictor of total tick counts p = 1.73e-8
  Plot was a significant predictor of total tick counts = p < 2e-16
  Year was a significant predictor of total tick counts p < 2e-16

*2.H1downB (negative binomial)*
  Site Intermidate was a significant predictor of total tick counts p = 2e-16
  Site Mesic was a significant predictor of total tick counts p = 4.59e-5

  Plot Partial was a significant predictor of total tick counts p = 3.37e-13
  Plot Total was a significant predictor of total tick counts p < 2e-16

  Year was a significant predictor of total tick counts p < 2e-16

*3. ZIP_round_downmam*
  Site was a significant predictor of total mammal counts p = 1.73e-8
  Plot was a significant predictor of total mammal counts = p < 2e-16
  Year was a significant predictor of total mammal counts p < 2e-16

*4.H1downB (negative binomial)*
  Site Intermidate was a significant predictor of total mammal counts p = 2e-16
  Site Mesic was a significant predictor of total mammal counts p = 0.00824

  Plot Partial was a significant predictor of total mammal counts p = 0.02398
  

  Year was a significant predictor of total mammal counts p < 2e-16
```{r zero inflated models}
library(pscl)
library(lmtest)

## check to make sure data really aren't normal 

# ticks forsure not normally distributed
length(combo$total_ticks[combo$total_ticks == 0]) # 879
shapiro.test(combo$total_ticks) # p < 2.2e-16


# mammals are more evelyn distributed, but this dataset may not be the appropriate way to run these analyses
length(combo$total_mammal[combo$total_mammal == 0]) # 6
shapiro.test(combo$total_mammal) # p < 2.2e-16

hist(combo$total_mammal)
min(combo$total_mammal, na.rm = TRUE) # 8
max(combo$total_mammal, na.rm = TRUE) # 72
length(combo$total_mammal) #1456
length(combo$total_mammal[combo$total_mammal == 0])/length(combo$total_mammal) #0.004120879
hist(combo$total_mammal, 50)
mean(combo$total_mammal, na.rm = TRUE) # 34.9131


######################## Tick abundances ##################

## part 1 of 0s
# zero inflated poisson (ZIP)
fldown <- formula(total_ticks ~ site + plot | site + plot) # what explains the load that's not 0 | explanatory variables if something is present or not
ZIP_round_down <- zeroinfl(fldown, dist = 'poisson',
                           link = 'logit',
                           data = combo)
summary(ZIP_round_down) # site and plot are signficant!

# zero inflated negative binomial (ZINB)
fldown <- formula(total_ticks ~ site + plot | site)
ZINB_round_down <- zeroinfl(fldown, dist = 'negbin', # changed distribution
                            link = 'logit',
                            data = combo)
summary(ZINB_round_down) # site and plot are significant!


# likeligood ratio test to compare ZIP and ZINB
# test the two models
lrtest(ZIP_round_down, ZINB_round_down) # ZINB p < 2.2e-16, so go with negative binomial


## part two non-0s
# Hurdle Models
# ZIP
H1downA <- hurdle(fldown, dist = "poisson", link = "logit",data = combo)
# ZINB
H1downB <- hurdle(fldown, dist = "negbin", link = "logit",data = combo)
lrtest(H1downA,H1downB) # Negative binomial version much more significant, p < 2.2e-16

summary(H1downB)



######################## Mammal abundances ########################

# zero inflated poisson
fldownmam <- formula(total_ticks ~ site + plot + total_mammal | year)
ZIP_round_downmam <- zeroinfl(fldown, dist = 'poisson',
                           link = 'logit',
                           data = combo)
summary(ZIP_round_downmam)

# zero adjusted poisson and negative binomial (hurdle models)
H1downAmam <- hurdle(fldownmam, dist = "poisson", link = "logit",data = combo)
H1downBmam <- hurdle(fldownmam, dist = "negbin", link = "logit",data = combo)
lrtest(H1downAmam,H1downBmam)
# Negative binomial version much more significant

summary(H1downBmam)
library(jtools)

```

Is there correlation between tick abundance and mammal abundance?
- even log transformed total ticks does not help normality, so will run a Spearman's rank correlation (non-parametric)

Spearman's rank correlation 
1. observations are random samples
2. relationship between X and Y are monotonic

```{r correlation}
shapiro.test(combo$total_ticks) # total ticks are not normally distributed, p < 2.2e-16

# breaks normality assumption so need to run a non-parametric test (i.e. Spearmans' rank)

# select only numeric values
combo.numeric <- data.frame(combo[,6], combo[,11:13])
colnames(combo.numeric) <- c('total_ticks', 'total_mammal', 'shannon', 'richness')

# if I wanted to a parametric test, I would try to log transform data to make it more normal
#combo.numeric$logticks <- log(combo.numeric$total_ticks + 1)

# visualize data
pairs.panels(combo.numeric, density = TRUE, cor = TRUE, lm = TRUE, method = 'spearman')

# run spearman's rank test

# total ticks & total mammals
cor.test(combo.numeric$total_ticks, combo.numeric$total_mammal,
         method = 'spearman') # rho = 0.007601 very very weak positive correlation

# mammal shannon diversity & total mammals
cor.test(combo.numeric$shannon, combo.numeric$total_mammal,
         method = 'spearman') # rho = 0.770606 very strong positive correlation but I think these are too colinear of variables

# mammal species richness & total mammals
cor.test(combo.numeric$richness, combo.numeric$total_mammal,
         method = 'spearman') # rho = 0.8226635 very strong positive correlation but I think these are too colinear of variables


```
X^2 goodness-of-fit test
- compares counts to a discrete probability distribution
H0: data comes from particular discrete probability distribution

- goodness-of-fit test compares count data to a model of the expected frequencies of a set of categories
H0: prob of tick presence occurring on any given month is equal to other proportions

assumptions:
- no more than 20% of categories have expected <5
- no category with expected < 1
```{r goodness of fit test}

```


```{r print out models}
library(MASS)
summary(glm.nb(total_ticks ~ site + plot + year, data = combo))

summary(glm(total_ticks ~ site + plot + year, data = combo,
    family = 'poisson'))



# library(broom.mixed)
# library(ggstance)
# 
# 
# summ(tick24)
# summ(mammal17) 
# summ(mammal14)
# summ(combo12) 
# summ(combo28)
# 
# plot_summs(combo1, scale = TRUE, plot.distributions = TRUE, inner_ci_level = .9)
# plot_summs(combo1, combo2, combo3)
# jtools::plot_summs(combo1)
# plot_summs(combo1)
# plot_summs(fit1, fit2, scale = TRUE, model.names = c('site', 'plot'))
# ## export this out to word
# install.packages('huxtable')
# library(huxtable)
# export_summs(fit1, fit2, scale = TRUE)
# 
# ## with confidence intervals instead of standard errors
# export_summs(fit1, fit2, scale = TRUE,
#             error_format = "[{conf.low}, {conf.high}]")
# 
# ##  save as word doc
# export_summs(fit1, fit2, scale = TRUE, 
#              to.file = "docx",
#              file.name = "test.docx")
```

