---
title: "Tejon_temperature_10222020"
author: "sbsambado"
date: "10/22/2020"
output: html_document
---


Import libraries and variables
```{r libraries and saved variables}
library(lubridate)
library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plyr)
library(car)
library(multcomp)
library(multcompView)
library(reshape2)
library(stringr)
library(car)
library(psych)
library(knitr)
library(nlme)
library(lme4)
library(kableExtra)
library(MASS)
library(grid)
library(vegan)
library(devtools)
library(ggfortify)
library(jtools)
library(effects)
library(pscl)
library(lmtest)
devtools::install_github('gavinsimpson/ggvegan')


clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('red4','red3','red')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))
xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))
```


Import data

```{r}
# full data set with just climate variables using hobo logger
hobo <- read.csv("~/Desktop/tejon_hobologger.csv")

hobo <- hobo[which(hobo$temp_F != "NA"),]
hobo <- hobo[which(hobo$rh_. != "NA"),]
hobo <- hobo[which(hobo$dewpt_F != "NA"),]

str(hobo)

# lubricated data set
hobo$date <-ymd(hobo$date)
str(hobo)

# creating new variable that incorporates date-time
Hobo <- mutate(hobo, fulldate = paste(date, time, sep = '-'))

Hobo$fulldate <- ymd_hms(Hobo$fulldate)
str(Hobo)

Hobo <- Hobo[which(Hobo$temp_F != "NA"),]
Hobo <- Hobo[which(Hobo$rh_. != "NA"),]
Hobo <- Hobo[which(Hobo$dewpt_F != "NA"),]


#### full data with climate and biotic using hobo logger

full <- read.csv("~/Desktop/tejon_tick_climate_10162020.csv")
#View(full)

full <- subset(full, select = -c(21:29))
full$Month <- factor(full$Month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
full$Site <- factor(full$Site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))
names(full)
names(full)[1] <- "year"
names(full)[2] <- "month"
names(full)[3] <- "site"
names(full)[4] <- "block"
names(full)[5] <- "plot"
names(full)[6] <- "plotID"
names(full)[7] <- "totalticks"

names(full)[15] <- "maxF"
names(full)[16] <- "minF"
names(full)[19] <- "precip_in"
names(full)[20] <- "precip_mm"


full$year <- factor(full$year)
full$stationID_elevation <- factor(full$stationID_elevation)

str(full)

Full <- full[which(full$maxF != "NA"),]
FULL <- full[which(full$rh_. != "NA"),]
FULL <- full[which(full$dewpt_F != "NA"),]

max(Full$maxF)

# make date column in Full
FULL <- mutate(Full, fulldate = paste(month, year, sep = '-'))

```


some plots
```{r}

HOBO <- hobo[which(hobo$temp_F < 95.0),]
ggplot(HOBO, aes(x = date, y = temp_F))+
  geom_line(aes(color = climate))+
  facet_wrap(~climate)+
  theme_bw() +
    geom_smooth(color = 'black')+
  theme(legend.position = 'bottom', legend.box = 'horizontal')+
  labs(x = 'Date', y = 'Temperature (F)', color = 'Site') +
  theme(strip.background = element_rect(fill = 'gray95')) +
  guides(color = FALSE)+
  geom_vline(xintercept= max(HOBO$temp_F),color='red')


ggplot(hobo, aes(x = date, y = dewpt_F))+
  geom_line(aes(color = climate))+
  facet_wrap(~climate)+
  theme_bw() +
  geom_smooth(color = 'black')+
  theme(legend.position = 'bottom', legend.box = 'horizontal')+
    labs(x = 'Date', y = 'Dewpoint', color = 'Site')+
  theme(strip.background = element_rect(fill = 'gray95')) +
  guides(color = FALSE)+
  geom_vline(xintercept= max(hobo$dewpt_F),color='red')



ggplot(hobo, aes(x = date, y = rh_.))+
  geom_line(aes(color = climate))+
  facet_wrap(~climate)+
  theme_bw() +
    geom_smooth(color = 'black')+
  theme(legend.position = 'bottom', legend.box = 'horizontal') +
  labs(x = 'Date', y = 'Relative Humidity', color = 'Site')+
  theme(strip.background = element_rect(fill = 'gray95')) +
  guides(color = FALSE)+
  geom_vline(xintercept= max(hobo$rh_.),color='red', size = 5)


# max(unique(Hobo$temp_F))
# ggplot(Hobo, aes(x = as.Date(fulldate), y = temp_F), group = climate, color = climate)+
#   geom_point(aes(color = climate), alpha = .5)+
#   facet_wrap(~climate)+
#   geom_line(aes(color = climate))+
#   theme(legend.position = 'bottom', legend.box = 'horizontal')+
#   scale_x_date(date_minor_breaks = "1 day")


ggplot(FULL, aes(x = month, y = maxF))+
  geom_point()+
  geom_line(aes(color = site))+
  facet_wrap(~site, ncol = 1)+
  theme_bw() +
    geom_smooth(color = 'black')+
  theme(legend.position = 'bottom', legend.box = 'horizontal')+
  labs(x = 'Date', y = 'Max Temperature (F)', color = 'Site') +
  theme(strip.background = element_rect(fill = 'gray95')) +
  guides(color = FALSE)+
  rotatexaxistheme


```
```{r random code for kurskal wallis and boxplot in ggpu r}
# Kurskal-Wallis test
# ```{r}
# kw1 <- kruskal.test(log_total ~ site, data = Tick)  
# # Kruskal-Wallis chi-squared = 77.186, df = 2, p-value < 2.2e-16
# 
# kw2 <- kruskal.test(log_total ~ plot, data = Tick) # p = 2.601e-12
# # Kruskal-Wallis chi-squared = 30.565, df = 2, p-value = 2.306e-07
# 
# library(rstatix)
# library(GmAMisc)
# kwPlot(Tick$log_total,Tick$site, strip = TRUE, posthoc = TRUE, adjust = "bonferroni")
# 
# Kw1 <- Tick %>% add_xy_position(x = 'site')
# ggboxplot(Tick, x = 'site', y = 'log_total')+
#   stat_pvalue_manual(Tick, hide.ns = TRUE) +
#   labs(subtitle = get_test_label(res.kruksal, detailed = TRUE),
#        caption = get_pwc_label(pwc))
# 
# library("ggpubr")
# ggboxplot(Tick, x = "site", y = "log_total", 
#           add = c("mean_se", "jitter"),
#           color = "site", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
#           order = c("Arid", "Intermediate", "Mesic"),
#           ylab = "Weight", xlab = "Treatment")
```

