---
title: "TejonAnalysis_02022021"
author: "sbsambado"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plyr)
library(car)
library(multcomp)
library(multcompView)
library(reshape2)
library(stringr)
library(car)
library(psych)
library(knitr)
library(nlme)
library(lme4)
library(kableExtra)
library(MASS)
library(grid)
library(vegan)
library(devtools)
library(ggfortify)
library(jtools)
library(effects)
library(pscl)
library(lmtest)
devtools::install_github('gavinsimpson/ggvegan')


clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('red4','red2','red')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))
xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))
```


Using data Steph sent on 11/16/2020
"Total_Ticks_2016_to_2018.csv"
"Tick_19.csv"

Merge 2 tick data sheets into 1 tick data sheet
```{r}

threeyrs <- read_csv("Total_Ticks_2016_to_2018.csv")
dim(threeyrs)
oneyr <- read_csv("Tick_19.csv")
dim(oneyr)
# get rid of survery column in threeyr since it wasn't completed in oneyrs 
## Maybe double check survey number doesn't matter
threeyrs <- subset(threeyrs, select = -c(3))

# make sure names are the same
## 2016 - 2018
names(threeyrs) <- tolower(names(threeyrs))

# change column names
names(threeyrs)[6] <- "total"
names(threeyrs)[7] <- "deoc"
names(threeyrs)[8] <- "ipac"
names(threeyrs)[9] <- "deva"
names(threeyrs)[10] <- "other"

## 2019
# make sure names are the same
names(oneyr) <- tolower(names(oneyr))

# change column names
names(oneyr)[6] <- "total"
names(oneyr)[7] <- "deoc"
names(oneyr)[8] <- "ipac"
names(oneyr)[9] <- "deva"
names(oneyr)[10] <- "other"


full <- rbind(threeyrs, oneyr)
dim(full) # 1449 10
```

Merge temp to tick data
```{r}
# merge temp data to full
temp <- read_csv("Tejon_temperature_02022021.csv")
Partial1 <- merge(full, temp)
dim(Partial1) # 1449 13
#View(Full)
```

Merge precip data to tick data
+ subset data came from tejon_tick_climate_10162020
```{r}
# precip <- read_csv("Tejon_precipitation_02022021.csv")
# Partial2 <- merge(Partial1, precip)
# dim(Partial2)
```

Merge mammal data to tick data
```{r}
# I calculated shannon diversity, species richness for each site/plot and assigned it a value
# mammal <- read_csv("tejon_mammal_02022021.csv")
# Partial3 <- merge(Partial2, mammal)
# dim(Partial3) 

```

Add metadata
```{r}
Full <- Partial1

##add station ID column
Full$station <- Full$site

  # assign station ID
  Full$station[Full$site == "Arid"] <- "Lebec"
  Full$station[Full$site == "Intermediate"] <- "Loraine"
  Full$station[Full$site == "Mesic"] <- "Chuchupate"
  
## add elevation
  Full$elevation <- Full$station
  Full$elevation[Full$station == "Lebec"] <- 1089.7
  Full$elevation[Full$station == "Loraine"] <- 1290.8
  Full$elevation[Full$station == "Chuchupate"] <- 1493.5

## add rain year
  # this is a dumb way but it is what it is
  Full$rain_year <- as.numeric(Full$year)
  Full$rain_year[Full$year == 2016 & Full$month == 'April'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'May'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'June'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'July'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'August'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'September'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'October'] <- "2015/2016"
  
  Full$rain_year[Full$year == 2016 & Full$month == 'November'] <- "2016/2017"
  Full$rain_year[Full$year == 2016 & Full$month == 'December'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'January'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'February'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'March'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'April'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'May'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'June'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'July'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'August'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'September'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'October'] <- "2016/2017"
  
  Full$rain_year[Full$year == 2017 & Full$month == 'November'] <- "2017/2018"
  Full$rain_year[Full$year == 2017 & Full$month == 'December'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'January'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'February'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'March'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'April'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'May'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'June'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'July'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'August'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'September'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'October'] <- "2017/2018"
  
  Full$rain_year[Full$year == 2018 & Full$month == 'November'] <- "2018/2019"
  Full$rain_year[Full$year == 2018 & Full$month == 'December'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'January'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'February'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'March'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'April'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'May'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'June'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'July'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'August'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'September'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'October'] <- "2018/2019"
  
  Full$rain_year[Full$year == 2019 & Full$month == 'November'] <- "2019/2020"
  Full$rain_year[Full$year == 2019 & Full$month == 'December'] <- "2019/2020"

## give plotID
Full <- transform(Full, plotID = as.numeric(interaction(block, plot, site, drop = TRUE)))
## create temp difference

Full$temp_diff <- Full$mean_maxF - Full$mean_minF
View(Full)

dim(Full) # 1449   17
```

give factor levels 
```{r}

Full$month <- factor(Full$month, 
                        levels = c("January", "February", "March",
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Full$site <- factor(Full$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))
```


aggregated tick counts + plots
```{r}
tick_month <- aggregate(total ~ month, data = Full, FUN = sum)
tick_plot <- aggregate(total ~ plot, data = Full, FUN = sum)
tick_site <- aggregate(total ~ site, data = Full, FUN = sum)

ggplot(tick_month, aes(x = month, y = total))+
  geom_histogram(stat = "identity")+
  theme_bw()+
  rotate_x_text() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")
  

ggplot(tick_plot, aes(x = plot, y = total, fill = plot))+
  geom_histogram(stat = "identity")+
  theme_bw() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))

ggplot(tick_site, aes(x = site, y = total, fill = site))+
  geom_histogram(stat = "identity")+
  theme_bw() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")+
  scale_fill_manual(values =c('red4','red3','red'))
```

basic plots with non-zero data (don't show these)
```{r}
Full_no0 <- Full[which(Full$total > 0),]
ggplot(Full_no0, aes(x = month, y = log(total)))+
  geom_boxplot()+
  geom_jitter(alpha = .4)+
  facet_wrap(~site, nrow = 3)+
  theme_minimal()+
  rotate_x_text()


ggplot(Full_no0, aes(x = site, y = log(total)))+
  geom_boxplot()+
  geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()

ggplot(Full_no0, aes(x = plot, y = log(total)))+
  geom_boxplot()+
  facet_wrap(~site)+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()
  
ggplot(Full_no0, aes(x = rain_year, y = log(total)))+
  geom_boxplot()+
  facet_wrap(~site)+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()

ggplot(Full_no0, aes(x = rain_year, y = log(total), color = site))+
  geom_boxplot()+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()

```


Total ticks ~ site + plot
```{r}
ggplot(Full_no0, aes(x = plot, y = log(total), fill = plot))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by plot") +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  guides(fill = FALSE)

ggplot(Full_no0, aes(x = site, y = log(total), fill = site))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by site") +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('red4','red3','red'))+
  guides(fill = FALSE)


ggplot(Full_no0, aes(x = plot, y = log(total), fill = site))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks)\n2016 - 2019', fill = 'Site',
       title = "Total ticks by site and plot") +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('red4','red3','red'))+
  theme(legend.position = 'bottom', legend.box = 'horizontal')

ggplot(Full_no0, aes(x = site, y = log(total), fill = plot))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks)\n2016 - 2019', fill = 'Plot',
       title = "Total ticks by site and plot") +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = 'bottom', legend.box = 'horizontal')
```

Total ticks ~ year
```{r}
ggplot(Full_no0, aes(x = rain_year, y = log(total), group = rain_year))+
  geom_boxplot(fill = "grey95")+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Rain Year', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by rain year") +
  xtitletheme + ytitletheme 

ggplot(Full_no0, aes(x = rain_year, y = log(total), group = rain_year, fill = site))+
  facet_wrap(~site)+
  scale_fill_manual(values=c('red4','red3','red'))+
  geom_boxplot()+
  theme_bw() +
  labs(x = 'Rain Year', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by rain year") +
  xtitletheme + ytitletheme + rotate_x_text() +
  theme(legend.position = 'bottom', legend.box = 'horizontal')
```

Total ticks ~ temp
```{r}
max <- ggplot(Full_no0, aes(x = mean_maxF, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw()+
  labs(x = 'Mean Max F', y = 'log(Total ticks)\n2016 - 2019')

min <-ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Min F', y = 'log(Total ticks)\n2016 - 2019')


diff <-ggplot(Full_no0, aes(x = temp_diff, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Difference in F', y = 'log(Total ticks)\n2016 - 2019')


ggarrange(max, diff + rremove('ylab'), min)

(overlap <- ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
  stat_smooth(data = Full_no0,aes(x = mean_maxF,ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3, color = 'darkblue')+
  stat_smooth(data = Full_no0,aes(x = mean_minF,ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3, color = 'blue')+
  theme_bw()+
  labs(x = 'Mean Min/Max in F', y = 'log(Total ticks)\n2016 - 2019'))

#### look at site

(max_site <- ggplot(Full_no0, aes(x = mean_maxF, y = log(total)))+
  facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw()+
  labs(x = 'Mean Max F', y = 'log(Total ticks)\n2016 - 2019'))

(min_site <-ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
    facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Min F', y = 'log(Total ticks)\n2016 - 2019'))


(diff_site <-ggplot(Full_no0, aes(x = temp_diff, y = log(total)))+
    facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Difference in F', y = 'log(Total ticks)\n2016 - 2019'))

ggarrange(max_site, diff_site + rremove('ylab'), min_site)


```


Total ticks ~ precipitation
```{r}
# ggplot(Full_no0, aes(y = log(precip_mm), x = site))+
#   geom_histogram(stat = "identity")+
#   geom_jitter(alpha = .3)+
#   theme_bw()+
#   facet_wrap(~year)+
#   labs(y = 'Precipiation (mm)', x = 'Site')+
#   scale_fill_manual(values =c('red4','red3','red'))
```

## Stats

ANOVA
```{r}
site1 <- aov(log(total)~ site, data = Full_no0)
plot(site1)

# check residuals
res = site1$residuals
shapiro.test(res) # p << 0.05 so data not normal
#hist(res)
leveneTest(total ~ site, 
        data = Full_no0) # p << 0.05 so variances are not equal
summary(site1)
# Fvalue = 39.52
# p < 2e-16 , reject H0 and conclude that 1 site is significantly different

#Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(site1)
# let's do this in a visual way
Post_hoc <-glht(site1, linfct = mcp(site = 'Tukey'))
summary(Post_hoc)

confint(Post_hoc) # pairwise test
cld(Post_hoc) # arid & mesic = a ; intermediate = b , so intermediate is different than arid & mesic
plot(Post_hoc) # plot 95% CIs

#### ^^ means ot total tick is significantly different in intermediate than mesic and arid

ggplot(data = Full_no0,
       aes(x = site, y = total,
           fill = site))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('red4','red2','red'))+
  theme_classic()+
  labs(x = 'Site', y = 'log(Total Ticks)', title = 'Total ticks by site')+
  #theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 3.7,
           size = 9) +
  annotate('text', label = 'b', 
           x = 2, y = 8.5,
           size = 9) +
    annotate('text', label = 'a', 
           x = 3, y = 4.1,
           size = 9)



## plot 

Full_no0$plot <- as.factor(Full_no0$plot)
plot1 <- aov(log(total)~ plot, data = Full_no0)
plot(site1)

# check residuals
res = plot1$residuals
shapiro.test(res) # p << 0.05 so data not normal
#hist(res)
leveneTest(total ~ plot, 
        data = Full_no0) # p << 0.05 so variances are not equal
summary(site1)
# Fvalue = 39.52
# p < 2e-16 , reject H0 and conclude that 1 site is significantly different

#Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(site1)
# let's do this in a visual way
Post_hoc <-glht(plot1, linfct = mcp(plot = 'Tukey'))
summary(Post_hoc)

confint(Post_hoc) # pairwise test
cld(Post_hoc) # arid & mesic = a ; intermediate = b , so intermediate is different than arid & mesic
plot(Post_hoc) # plot 95% CIs

#### ^^ means ot total tick is significantly different in intermediate than mesic and arid

ggplot(data = Full_no0,
       aes(x = plot, y = total,
           fill = plot))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme_classic()+
  labs(x = 'Site', y = 'log(Total Ticks)', title = 'Total ticks by plot')+
  #theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 4.3,
           size = 9) +
  annotate('text', label = 'b', 
           x = 2, y = 5.9,
           size = 9) +
    annotate('text', label = 'b', 
           x = 3, y = 7.1,
           size = 9)




```

GLMS

 #**Generalized Linear Models**
#What are good predictors (X) of total tick abundance (Y)?
**Ran *generalized linear models* because data breaks assumptions of normality. Used family poisson because of count data.**
*Object **'tickmodels'** is table of df, AIC, BIC, variables used in glm models*
+ *lowest AIC: all 4 variables (Site, Year, Plot, Month)*
+ *lowest AIC for a single predictor = Year*
  + **Year AIC = 10496**
  + **Site AIC = 10665**
  + **Plot AIC = 10871**
models : tick1 - tick7 

```{r GLM for tick data}
tick1 <- glm(total ~ rain_year, 
               data = Full_no0,
             family = 'poisson')
tick2 <- glm(total ~ site, 
               data = Full_no0,
              family = 'poisson')
tick3 <- glm(total ~ plot, 
               data = Full_no0,
              family = 'poisson')
tick4 <- glm(total ~ month, 
               data = Full_no0,
              family = 'poisson')
tick5 <- glm(total ~ site + rain_year, 
               data = Full_no0,
              family = 'poisson')
tick6 <- glm(total ~ site + rain_year + plot, 
               data = Full_no0,
              family = 'poisson')
tick7 <- glm(total ~ site + rain_year + plot + month, 
               data = Full_no0,
              family = 'poisson')
# calculate AIC of each model
tickresults <- AIC(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7)
# add other metrics to table
tickmodels <- list(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7) # keep models in the same order as they were created
# BIC places higher penalty on extra parameter
tickresults$BIC <- sapply(tickmodels, BIC) # add column for BIC
tick_model_summary <- lapply(tickmodels, summary)
# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(tickmodels)) {
  tickresults$rsq[i] <- tick_model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  tickresults$adj_rsq[i] <- tick_model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again
# add column with variables used in model
tickresults$variables <- c('rain_year','site', 'plot', 'month', 'site + rain_year', 'site + rain_year + plot', 'site + rain_year + plot + month')
# make a nice table!
tickmodels <-kable(tickresults, digits = 2, align = 'c') %>%
  kable_styling()
eighteen <-tickmodels
eighteen
#nineteen <- export_summs(tick1, tick2, tick3, scale = TRUE)
#nineteen
#min(tickresults$AIC)

```


Negative binomial glms

```{r}
# not normal
hist(log(Full_no0$total + 1))
hist(log(Full$total+ 1))

glm.negbin = glm.nb(log(total) ~ site,data = Full_no0)
summary(glm.negbin)

glm.fit = glm(log(total)~site, data = Full_no0,
              family = "poisson")
summary(glm.fit)

glm.quasipoisson = update(glm.fit, family = quasipoisson)
summary(glm.quasipoisson)

# hillary's suggestions
glm1 <- glmer(total ~ mean_minF + plot +  mean_minF*plot + ( 1 | plotID) + (1 | month) + (1 |rain_year),
              data = Full_no0, family = poisson)
summary(glm1)
anova(glm1)
plot(predictorEffects(glm1),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))

```

Full models
```{r}
tick_climate.Full_no0 <- glm.nb(total~site*plot*temp_diff, data=Full_no0)
tick_climate.Full_no0_nointeraction <- glm.nb(total~site+plot+temp_diff, data=Full_no0) # lowest BIC score
tick_climate.S <- glm.nb(total~site, data=Full_no0)
tick_climate.M <- glm.nb(total~plot, data=Full_no0)
tick_climate.D <- glm.nb(total~temp_diff, data=Full_no0)
tick_climate.SplusM <- glm.nb(total~site+plot, data=Full_no0)
tick_climate.StimesM <- glm.nb(total~site*plot, data=Full_no0)
tick_climate.SplusD <- glm.nb(total~site+temp_diff, data=Full_no0)
tick_climate.StimesD <- glm.nb(total~site*temp_diff, data=Full_no0)
tick_climate.MplusD <- glm.nb(total~plot+temp_diff , data=Full_no0)
tick_climate.MtimesD <- glm.nb(total~plot*temp_diff, data=Full_no0)
tick_climate.SplusMD <- glm.nb(total~site+plot*temp_diff, data=Full_no0)
tick_climate.MplusSD <- glm.nb(total~plot+site*temp_diff, data=Full_no0)
tick_climate.DplusSM <- glm.nb(total~site*plot+temp_diff, data=Full_no0)
AIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
BIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
temp_diff <- BIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
min_dif <-min(temp_diff$BIC)
plot(predictorEffects(tick_climate.Full_no0, ~ plot),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
plot(predictorEffects(tick_climate.Full_no0, ~ site),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
plot(predictorEffects(tick_climate.Full_no0, ~ temp_diff),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
```

