---
title: "Tejon figure code"
author: "sbsambado"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(xlsx)
library(dplyr)
library(ggpubr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plyr)
library(car)
library(multcomp)
library(multcompView)
library(reshape2)
library(stringr)
library(car)
library(psych)
library(knitr)
library(nlme)
library(lme4)
library(kableExtra)
library(MASS)
library(grid)
library(vegan)
library(devtools)
library(ggfortify)
library(jtools)
library(effects)
library(pscl)
library(lmtest)
library(ggstance)
library(huxtable)
library(qwraps2)
library(gtsummary)
library(Rcapture)
library(xlsx)
library(glmmTMB) 
library(tidyverse) 
library(gt) 
library(patchwork) 
library(effects) 
library(jtools)

clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('brown1','tan1','lightgoldenrod1')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))
xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))


```

### Dataset for tick mean analyses

Merge Data and tidy it up
```{r message = FALSE}

threeyrs <- read_csv("Total_Ticks_2016_to_2018.csv")
dim(threeyrs)
oneyr <- read_csv("Tick_19.csv")
dim(oneyr)
# get rid of survery column in threeyr since it wasn't completed in oneyrs 
## Maybe double check survey number doesn't matter
threeyrs <- subset(threeyrs, select = -c(3))

# make sure names are the same
## 2016 - 2018
names(threeyrs) <- tolower(names(threeyrs))

# change column names
names(threeyrs)[6] <- "total"
names(threeyrs)[7] <- "deoc"
names(threeyrs)[8] <- "ipac"
names(threeyrs)[9] <- "deva"
names(threeyrs)[10] <- "other"

## 2019
# make sure names are the same
names(oneyr) <- tolower(names(oneyr))

# change column names
names(oneyr)[6] <- "total"
names(oneyr)[7] <- "deoc"
names(oneyr)[8] <- "ipac"
names(oneyr)[9] <- "deva"
names(oneyr)[10] <- "other"


tick <- rbind(threeyrs, oneyr)

# log transform data
tick$log_total <- log(tick$total + 1)

## FINAL DATASET TO COMPARE TICK MEANS
dim(tick) # 1449 11
str(tick)

tick$year <- as.numeric(tick$year)
tick$site <- as.factor(tick$site)
tick$plot <- as.factor(tick$plot)
Tick <- tick
```

### Dataset for tick density relationships

Lots of movin' and groovin here. Trying to add multiple datasets together while avoiding duplication and too much observation lost

Merge Data
```{r building big dataset with individual datasets, message=FALSE}
##################################################################################################
# TICK COUNTS (YEAR, MONTH, PLOT, SITE)
#################################################################################################

# use cleaned up version that was used for mean tick count analyses = Tick
dim(Tick) # 1449 observations, 11 rows


##################################################################################################
# CLIMATE (TEMP, PRECIP)
#################################################################################################

############## temperature ##############
temp <- read_csv("Tejon_temperature_02022021.csv")
temp.tick <- merge(Tick, temp)
dim(temp.tick) # 1449, 14 (added meanmaxF, meanminF, station)
#View(temp.tick)

############## precipitation ##############

precip <- read_csv("tejon_tick_climate_10162020.csv")

precip_subset <- subset(precip, select = c(1:5,19,20))
names(precip_subset) <- tolower(names(precip_subset))
temp.tick.precip <- join(temp.tick, precip_subset)
dim(temp.tick.precip) #2025, 16 (added precip (in), precip (mm))
names(temp.tick.precip)
# seems like data got duplicated let's fix that

# best way is to remove duplicate rows which I lose ~50 observations but I think that is a better way to do it then extract unique elements by columns since alot of these observations are repetitive

temp.tick.prep.short <- temp.tick.precip %>%
  distinct()

dim(temp.tick.prep.short) # 1395, 16

##################################################################################################
# VERTEBRATE (SPIECES RICHNESS, DIVERSITY, POP ESTIMATES)
#################################################################################################

############## mammal metrics ############## 

mam <- read_csv("tejon_mammal_02022021.csv")
names(mam) <- tolower(names(mam))

climate.mam <- join(temp.tick.prep.short, mam)
dim(climate.mam) # 1895, 19 (added cum_mammal, shan_mammal, rich_mammal)

# remove duplicate rows

climate.mam.short <- climate.mam %>%
  distinct()
dim(climate.mam.short) #1395, 19 # same amount of observations after climate dataset


############## pop estimates ############## 

pop_estimates <-read_csv("tejon_mamliz_popestimates.csv")
dim(pop_estimates) #27 5

climate.mam.combo <- join(climate.mam.short, pop_estimates)
dim(climate.mam.combo) # 1395   21 (added liz estimates, mam estimates)
names(climate.mam.combo)[20] <- "liz_estimates"

dim(climate.mam.combo) # 1395   21
names(climate.mam.combo)

```

Tidy data
```{r tidy final big data}
# let's shape this final sheet

## add some more metadata
# add Celcius
climate.mam.combo$mean_maxC <- (climate.mam.combo$mean_maxF - 32)*(5/9)
climate.mam.combo$mean_minC <- (climate.mam.combo$mean_minF - 32)*(5/9)


# add temperature difference 
climate.mam.combo$tempdifF <- (climate.mam.combo$mean_maxF - climate.mam.combo$mean_minF)
climate.mam.combo$tempdifC <- (climate.mam.combo$mean_maxC - climate.mam.combo$mean_minC)



# add elevation
  climate.mam.combo$elevation <- climate.mam.combo$station
  climate.mam.combo$elevation[climate.mam.combo$station == "Lebec"] <- 1089.7
  climate.mam.combo$elevation[climate.mam.combo$station == "Loraine"] <- 1290.8
  climate.mam.combo$elevation[climate.mam.combo$station == "Chuchupate"] <- 1493.5

# add rain year
  # this is a dumb way but it is what it is
  climate.mam.combo$rain_year <- as.numeric(climate.mam.combo$year)
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'April'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'May'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'June'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'July'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'August'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'September'] <- "2015/2016"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'October'] <- "2015/2016"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'November'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2016 & climate.mam.combo$month == 'December'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'January'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'February'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'March'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'April'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'May'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'June'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'July'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'August'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'September'] <- "2016/2017"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'October'] <- "2016/2017"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'November'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2017 & climate.mam.combo$month == 'December'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'January'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'February'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'March'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'April'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'May'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'June'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'July'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'August'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'September'] <- "2017/2018"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'October'] <- "2017/2018"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'November'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2018 & climate.mam.combo$month == 'December'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'January'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'February'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'March'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'April'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'May'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'June'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'July'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'August'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'September'] <- "2018/2019"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'October'] <- "2018/2019"
  
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'November'] <- "2019/2020"
  climate.mam.combo$rain_year[climate.mam.combo$year == 2019 & climate.mam.combo$month == 'December'] <- "2019/2020"

# add plotID
climate.mam.combo <- transform(climate.mam.combo, plotidname = as.character(interaction(site, plot,block, drop = TRUE)))
climate.mam.combo <- transform(climate.mam.combo, plotID = as.numeric(interaction(site, plot,block, drop = TRUE)))
str(climate.mam.combo) # 1395 observations, 26 variables (added plotidname, plotID)

# put data in right format
str(climate.mam.combo)

climate.mam.combo$year <- as.factor(climate.mam.combo$year)
climate.mam.combo$block <- as.factor(climate.mam.combo$block)
climate.mam.combo$plotID <- as.factor(climate.mam.combo$plotID)
climate.mam.combo$rain_year <- as.factor(climate.mam.combo$rain_year)
climate.mam.combo$elevation <- as.numeric(climate.mam.combo$elevation)

climate.mam.combo$month <- factor(climate.mam.combo$month, 
                        levels = c('January', 'February', 'March',
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
climate.mam.combo$site <- factor(climate.mam.combo$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))

climate.mam.combo$plot <- factor(climate.mam.combo$plot, 
                    levels = c('Open', 'Partial', 'Total'))

# rename precip columns
names(climate.mam.combo)[15] <- "precip_in"
names(climate.mam.combo)[16] <- "precip_mm"


# Finally, the final dataset for GLMMS
FINAL <- climate.mam.combo
str(FINAL) # 1395 observations, 30 variables

# write to an excel sheet
write.xlsx(FINAL, file = "Tejon_MixedModels_Dataset.xlsx",
           col.names = TRUE, row.names = TRUE)
```

### Data figures

**Fig 1. Tick means: One variable plots (plot or site)**
```{r}

# omit 0s to make plots look nicer, but this may not be kosher
Full_no0 <- FINAL[which(FINAL$log_total != 0),]

plot1 <- ggplot(Full_no0, aes(x = plot, y = log_total, fill = plot))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  guides(fill = FALSE)+
  annotate('text', label='A', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

site1 <-ggplot(Full_no0, aes(x = site, y = log_total, fill = site))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Site', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)+
  annotate('text', label='B', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

var_plot <-ggarrange(plot1 + rremove('ylab'), site1 + rremove('ylab'),
          ncol = 2)
annotate_figure(var_plot,
          left = text_grob("Log of total ticks (2016 - 2019)", rot = 90, face = "bold"))
ggsave("Fig1_onevargraph.jpeg", dpi = 320)          
          

```


**Fig 2. Tick means: Bivariate plot**
```{r}

ggplot(Full_no0, aes(x = site, y = log_total, fill = plot))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Site', y = "Log of total ticks (2016 - 2019)", fill = 'Plot') +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = c(.068, .85), # horizontal, verticle
        legend.box.background = element_rect(color = 'black', fill =NA))
ggsave("Fig2_twovargraph.jpeg", dpi = 320)  
```


**Fig 3. Temperature: Difference in monthly temp**
```{r}


ggplot(Full_no0, aes(y = tempdifC, x = site, fill = site))+
  geom_violin(scale = 'count',trim = TRUE, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(alpha = .15, width = .2)+
  theme_bw()+
  labs(x = 'Site', y = 'Difference in monthly mean temperature (°C)', fill = "Site")+
  xtitletheme + ytitletheme +
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)
  #theme(legend.position = c(.093, .85),
        #legend.box.background = element_rect(color = 'black', fill =NA))
ggsave("Fig3_difftempCsite.jpeg", dpi = 320)

```


**Fig 4. Vertebrate populations: lizard and mammal population estimates**
```{r}
pop_estimates <-read_csv("tejon_mamliz_popestimates.csv")


liz <- ggplot(pop_estimates, aes(x = plot, y = liz_estimate, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  labs(x = "Plot", y = "Lizard Population Estimates")+
  guides(fill = FALSE)

mam<- ggplot(pop_estimates, aes(x = plot, y = mam_estimates, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4')) +
  labs(x = "Plot", y = "Mammal Population Estimates")+
  guides(fill = FALSE)

estimates <- ggarrange(liz, mam)

ggsave("PopulationEstimates.jpeg", dpi = 320)



### OR without individual blocks (need to go back to find old version of this)

# site <- c("Arid","Arid","Arid","Intermediate","Intermediate","Intermediate", "Mesic","Mesic","Mesic")
# plot <- c("Open", "Partial", "Total","Open", "Partial", "Total","Open", "Partial", "Total")
# mcounts <- c(7,85,72,10,18,38,64,142,71)
# 
# mcount_table <- cbind(site, plot, mcounts)
# 
# mcount_table <-as.data.frame(mcount_table)
# 
# mcount_table$vert <- mcount_table$mcounts
# mcount_table$vert <- "mammal"
# names(mcount_table) <- c("site", "plot", "count", "vert")
# 
# 
# count_table <- mcount_table
# count_table$vert <- count_table$lizardcounts
# count_table$vert <- "lizard"
# names(count_table) <- c("site", "plot", "count", "vert")
# 
# 
# estimate <- rbind(mcount_table, count_table)
# str(estimate)
# 
# estimate$count <- as.numeric(estimate$count)
# 
# ggplot(estimate, aes(x = plot, y = count, fill = vert, group_by(vert)))+
#   geom_boxplot()+
#   theme_bw()+
#   xtitletheme + ytitletheme +
#   scale_fill_manual(values=c('grey50', "grey90")) +
#   labs(x = "Plot", y = "Vertebrate Population Estimates", fill = "Vertebrate") +
#   theme(legend.position = "bottom")
# ggsave("PopulationEstimate_aggregated.jpeg", dpi = 320)
  



```

~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
SUPPLEMENTAL?
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

**Fig 5. Tick means: By month**
```{r}

ggplot(Full_no0, aes(x = site, y = log_total, fill = site))+
  facet_wrap(~ month)+
  geom_boxplot()+
  theme_bw() +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'),
        axis.text.x = element_text(angle = 45, size = 8, vjust = 0.5))+
  labs(x = 'Site', y = 'Log of total ticks (2016 - 2019)') +
  xtitletheme + ytitletheme +
  guides(fill = FALSE)+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))
ggsave("TickMeans_SiteMonth.jpeg", dpi = 320)

### OR ####
ggplot(Full_no0, aes(x = site, y = log_total, fill = plot))+
  facet_wrap(~ month)+
  geom_boxplot()+
  theme_bw() +
  theme(strip.background = element_rect(fill = "gray95"),
        strip.text = element_text(size=6, face = 'bold'),
        axis.text.x = element_text(angle = 45, size = 8, vjust = 0.5))+
  labs(x = 'Plot', y = 'Log of total ticks (2016 - 2019)') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = c(.81,.01), legend.direction = 'horizontal')
ggsave("TickMeans_PlotMonth.jpeg", dpi = 320)



```


More graphs

**Pop estimates**
```{r}
pop_estimates <-read_csv("tejon_mamliz_popestimates.csv")


(liz <- ggplot(pop_estimates, aes(x = plot, y = liz_estimate, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  labs(x = "Plot", y = "Lizard Population Estimates")+
  guides(fill = FALSE))

(mam<- ggplot(pop_estimates, aes(x = plot, y = mam_estimates, fill = plot))+
  geom_boxplot()+
  theme_bw()+
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4')) +
  labs(x = "Plot", y = "Mammal Population Estimates")+
  guides(fill = FALSE))

estimates <- ggarrange(liz, mam)
# 
#ggsave("PopulationEstimatesBlock.jpeg", dpi = 320)


```


**Estimates for 3 model types**
```{r}
plot_summs(host_gnb_mod4, climate_gnbmod5, combo_gnb_mod7,scale = TRUE,
           model.names = c("Host", "Climate", "Combination"))
```







`

