---
title: "TejonAnalysis_02022021"
author: "sbsambado"
date: "2/2/2021"
output: html_document
---

Fig 1: 2 plots (1 variable plot, 1 variable site)
Fig 2: 1 plot (2 variables of boxplot, maybe do x = site, plot = fill)
Fig 3: 1 plot (3 lines for site for difference in temp of mean)
Table 1: model outputs

Supplemental
-mammal
-lizard
-precip
-predict


Analyses
1. wilcox rank test for treatment vs control [medians total ticks]
  a. site (arid vs mesic)
  b. plote (open vs total)
  
2. ONEANOVA [means total ticks]
  a. 3 sites
  b. 3 plots

2b. TWO WAY ANOVA
```{r}
site_subset <- Full[which(Full$site == "Arid" | Full$site == "Mesic"),]
plot_subset <- Full[which(Full$plot == "Open" | Full$plot == "Total"),]


aov_subset <- site_subset[which(site_subset$plot == "Open" |site_subset$plot == "Total"), ]
av_model <- aov(total ~ plot + site, data = aov_subset)
summary(av_model) # both sig, plot more sig

qqPlot(av_model$residuals)
# plot 38.30 1.11e-09 ***
# site 24.45 9.90e-07 ***

## interaction term
av_model2 <- aov(total ~ plot + site + (plot*site), data = aov_subset)
summary(av_model2) # interaction not significant


TukeyHSD(av_model)

# statisitcal inference

## one way two way ANOVA check whether a numeric variable varies accdoring to the levels of 1 to 2 cateforical variables

# berfore prediction
```

3. Mixed models
- site, plot
- site, plot, rain year
- site, plot, mammal, precip, lizard, humidity

emmeans 
plot
fits

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(ggpubr)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(plyr)
library(car)
library(multcomp)
library(multcompView)
library(reshape2)
library(stringr)
library(car)
library(psych)
library(knitr)
library(nlme)
library(lme4)
library(kableExtra)
library(MASS)
library(grid)
library(vegan)
library(devtools)
library(ggfortify)
library(jtools)
library(effects)
library(pscl)
library(lmtest)
devtools::install_github('gavinsimpson/ggvegan')


clean_background <- theme(plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("white"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12),
        legend.key = element_rect("white"))

plotcolor = c('darkseagreen1','darkseagreen3','darkseagreen4')
sitecolor = c('brown1','tan1','lightgoldenrod1')

rotatexaxistheme <- theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ytitletheme <- theme(axis.title.y = element_text(face = 'bold', size = 12, vjust = 0.5))
xtitletheme <- theme(axis.title.x = element_text(face = 'bold', size = 12, vjust = 0.5))
```


Using data Steph sent on 11/16/2020
"Total_Ticks_2016_to_2018.csv"
"Tick_19.csv"

Merge 2 tick data sheets into 1 tick data sheet
```{r}

threeyrs <- read_csv("Total_Ticks_2016_to_2018.csv")
dim(threeyrs)
oneyr <- read_csv("Tick_19.csv")
dim(oneyr)
# get rid of survery column in threeyr since it wasn't completed in oneyrs 
## Maybe double check survey number doesn't matter
threeyrs <- subset(threeyrs, select = -c(3))

# make sure names are the same
## 2016 - 2018
names(threeyrs) <- tolower(names(threeyrs))

# change column names
names(threeyrs)[6] <- "total"
names(threeyrs)[7] <- "deoc"
names(threeyrs)[8] <- "ipac"
names(threeyrs)[9] <- "deva"
names(threeyrs)[10] <- "other"

## 2019
# make sure names are the same
names(oneyr) <- tolower(names(oneyr))

# change column names
names(oneyr)[6] <- "total"
names(oneyr)[7] <- "deoc"
names(oneyr)[8] <- "ipac"
names(oneyr)[9] <- "deva"
names(oneyr)[10] <- "other"


full <- rbind(threeyrs, oneyr)
dim(full) # 1449 10
```

Merge temp to tick data
```{r}
# merge temp data to full
temp <- read_csv("Tejon_temperature_02022021.csv")
Partial1 <- merge(full, temp)
dim(Partial1) # 1449 13
#View(Full)
```

Merge precip data to tick data
+ subset data came from tejon_tick_climate_10162020
```{r}
# precip <- read_csv("Tejon_precipitation_02022021.csv")
# Partial2 <- merge(Partial1, precip)
# dim(Partial2)
```

Merge mammal data to tick data
```{r}
# I calculated shannon diversity, species richness for each site/plot and assigned it a value
# mammal <- read_csv("tejon_mammal_02022021.csv")
# Partial3 <- merge(Partial2, mammal)
# dim(Partial3) 

```

Add metadata
```{r}
Full <- Partial1

##add station ID column
Full$station <- Full$site

  # assign station ID
  Full$station[Full$site == "Arid"] <- "Lebec"
  Full$station[Full$site == "Intermediate"] <- "Loraine"
  Full$station[Full$site == "Mesic"] <- "Chuchupate"
  
## add elevation
  Full$elevation <- Full$station
  Full$elevation[Full$station == "Lebec"] <- 1089.7
  Full$elevation[Full$station == "Loraine"] <- 1290.8
  Full$elevation[Full$station == "Chuchupate"] <- 1493.5

## add rain year
  # this is a dumb way but it is what it is
  Full$rain_year <- as.numeric(Full$year)
  Full$rain_year[Full$year == 2016 & Full$month == 'April'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'May'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'June'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'July'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'August'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'September'] <- "2015/2016"
  Full$rain_year[Full$year == 2016 & Full$month == 'October'] <- "2015/2016"
  
  Full$rain_year[Full$year == 2016 & Full$month == 'November'] <- "2016/2017"
  Full$rain_year[Full$year == 2016 & Full$month == 'December'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'January'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'February'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'March'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'April'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'May'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'June'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'July'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'August'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'September'] <- "2016/2017"
  Full$rain_year[Full$year == 2017 & Full$month == 'October'] <- "2016/2017"
  
  Full$rain_year[Full$year == 2017 & Full$month == 'November'] <- "2017/2018"
  Full$rain_year[Full$year == 2017 & Full$month == 'December'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'January'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'February'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'March'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'April'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'May'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'June'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'July'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'August'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'September'] <- "2017/2018"
  Full$rain_year[Full$year == 2018 & Full$month == 'October'] <- "2017/2018"
  
  Full$rain_year[Full$year == 2018 & Full$month == 'November'] <- "2018/2019"
  Full$rain_year[Full$year == 2018 & Full$month == 'December'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'January'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'February'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'March'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'April'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'May'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'June'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'July'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'August'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'September'] <- "2018/2019"
  Full$rain_year[Full$year == 2019 & Full$month == 'October'] <- "2018/2019"
  
  Full$rain_year[Full$year == 2019 & Full$month == 'November'] <- "2019/2020"
  Full$rain_year[Full$year == 2019 & Full$month == 'December'] <- "2019/2020"

## give plotID
Full <- transform(Full, plotID = as.numeric(interaction(block, plot, site, drop = TRUE)))
## create temp difference

Full$temp_diff <- Full$mean_maxF - Full$mean_minF
View(Full)

dim(Full) # 1449   17
```

Basic data stats
```{r}
summary(Full)
dim(Full) # 1449 observations, 17 variables

totalcount <- Full %>% 
  tally(total) # 2692 ticks

ipaccount <- Full %>% 
  tally(ipac) # 206 ticks 



deoccount <- Full %>% 
  tally(deoc) # 2420 ticks

devacount <- Full %>% 
  tally(deva) # 68 ticks

Full %>% 
  tally(other) # 2 ticks

plotcount <- Full %>% 
  group_by(plot) %>% 
  tally(total) # open = 488, partial = 976, total = 1228

sitecount <- Full %>% 
  group_by(site) %>% 
  tally(total) # arid = 340, intermediate = 1764, mesic = 588

# calculate % of each species

ipaccount/totalcount * 100 # 7.652303% of tick pop is ipac

deoccount/totalcount*100 # 89.89599%
devacount/totalcount*100 # 2.526003

(deoccount + devacount)/totalcount *100 # 92.42199% of tick pop is derm


```


give factor levels 
```{r}

Full$month <- factor(Full$month, 
                        levels = c("January", "February", "March",
                                   'April', 'May', 'June', 'July',
                                   'August', 'September', 'October',
                                   'November', 'December'))
Full$site <- factor(Full$site, 
                    levels = c('Arid', 'Intermediate', 'Mesic'))
```

aggregated tick counts + plots
```{r}
tick_month <- aggregate(total ~ month, data = Full, FUN = sum)
tick_plot <- aggregate(total ~ plot, data = Full, FUN = sum)
tick_site <- aggregate(total ~ site, data = Full, FUN = sum)

ggplot(tick_month, aes(x = month, y = total))+
  geom_histogram(stat = "identity")+
  theme_bw()+
  rotate_x_text() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")
  

ggplot(tick_plot, aes(x = plot, y = total, fill = plot))+
  geom_histogram(stat = "identity")+
  theme_bw() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")+
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))

ggplot(tick_site, aes(x = site, y = total, fill = site))+
  geom_histogram(stat = "identity")+
  theme_bw() +
  labs(x = "Month", y = "Aggregated tick counts\n (2016-2019)")+
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))
```

basic plots with non-zero data (don't show these)
```{r}
Full_no0 <- Full[which(Full$total > 0),]
ggplot(Full_no0, aes(x = month, y = log(total)))+
  geom_boxplot()+
  geom_jitter(alpha = .4)+
  facet_wrap(~site, nrow = 3)+
  theme_minimal()+
  rotate_x_text()


ggplot(Full_no0, aes(x = site, y = log(total)))+
  geom_boxplot()+
  geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()

ggplot(Full_no0, aes(x = plot, y = log(total)))+
  geom_boxplot()+
  facet_wrap(~site)+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()
  
ggplot(Full_no0, aes(x = rain_year, y = log(total)))+
  geom_boxplot()+
  facet_wrap(~site)+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()+
  rotate_x_text()

ggplot(Full_no0, aes(x = rain_year, y = log(total), color = site))+
  geom_boxplot()+
  #geom_jitter(alpha = .4, aes(color = site))+
  theme_minimal()

```


Total ticks ~ site + plot

Fig 1. one var plots
```{r}


plot1 <- ggplot(Full_no0, aes(x = plot, y = log(total), fill = plot))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  guides(fill = FALSE)+
  annotate('text', label='A', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

site1 <-ggplot(Full_no0, aes(x = site, y = log(total), fill = site))+
  geom_boxplot()+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Site', y = 'log(Total ticks) 2016 - 2019') +
  xtitletheme + ytitletheme +
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)+
  annotate('text', label='B', x=-Inf, y=Inf, hjust=-.5, vjust=2,size = 7)

var_plot <-ggarrange(plot1 + rremove('ylab'), site1 + rremove('ylab'),
          ncol = 2)
annotate_figure(var_plot,
          left = text_grob("Log of total ticks (2016 - 2019)", rot = 90, face = "bold"))
ggsave("Fig1_onevargraph.jpeg", dpi = 320)          
          

```


Fig 2. two var plots
```{r}
ggplot(Full_no0, aes(x = plot, y = log(total), fill = site))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Plot', y = 'log(Total ticks)\n2016 - 2019', fill = 'Site',
       title = "Total ticks by site and plot") +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  theme(legend.position = 'bottom', legend.box = 'horizontal')



ggplot(Full_no0, aes(x = site, y = log(total), fill = plot))+
  geom_boxplot()+
  geom_point(alpha = .15)+
  theme_bw() +
  labs(x = 'Site', y = "Log of total ticks (2016 - 2019)", fill = 'Plot') +
  xtitletheme + ytitletheme+
  scale_fill_manual(values =c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme(legend.position = c(.068, .85), # horizontal, verticle
        legend.box.background = element_rect(color = 'black', fill =NA))
ggsave("Fig2_twovargraph.jpeg", dpi = 320)  
```

Fig 3. temp diff by site
```{r}
ggplot(Full_no0, aes(x = temp_diff, y = total, group = site, color = site))+
  geom_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)),lwd = 2)+
  theme_bw()+
  labs(x = 'Difference in min/max (F)', y = 'Total ticks')+
  xtitletheme + ytitletheme +
  scale_color_manual(values =c('brown1','tan1','lightgoldenrod1'))
#ggsave("TemperatureDiffTick_v2.pdf", dpi = 320)

ggplot(Full_no0, aes(y = temp_diff, x = site, fill = site))+
  geom_violin(scale = 'count',trim = TRUE, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(alpha = .15, width = .2)+
  theme_bw()+
  labs(x = 'Site', y = 'Difference in monthly mean temperature (F)', fill = "Site")+
  xtitletheme + ytitletheme +
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  theme(legend.position = c(.1, .85),
        legend.box.background = element_rect(color = 'black', fill =NA)) # horizontal, verticle
ggsave("Fig3_difftempFsite.jpeg", dpi = 320)


# convert to C
Full_no0$mean_maxC <- (Full_no0$mean_maxF - 32)*(5/9)
Full_no0$mean_minC <- (Full_no0$mean_minF - 32)*(5/9)

Full_no0$temp_diffC <- Full_no0$mean_maxC  - Full_no0$mean_minC 

ggplot(Full_no0, aes(y = temp_diffC, x = site, fill = site))+
  geom_violin(scale = 'count',trim = TRUE, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(alpha = .15, width = .2)+
  theme_bw()+
  labs(x = 'Site', y = 'Difference in monthly mean temperature (°C)', fill = "Site")+
  xtitletheme + ytitletheme +
  scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))+
  guides(fill = FALSE)
  #theme(legend.position = c(.093, .85),
        #legend.box.background = element_rect(color = 'black', fill =NA))
ggsave("Fig3_difftempCsite.jpeg", dpi = 320)

```

Total ticks ~ year
```{r}
ggplot(Full_no0, aes(x = rain_year, y = log(total), group = rain_year))+
  geom_boxplot(fill = "grey95")+
  geom_jitter(alpha = .15, width = .2)+
  theme_bw() +
  labs(x = 'Rain Year', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by rain year") +
  xtitletheme + ytitletheme 

ggplot(Full_no0, aes(x = rain_year, y = log(total), group = rain_year, fill = site))+
  facet_wrap(~site)+
  scale_fill_manual(values=c('brown1','tan1','lightgoldenrod1'))+
  geom_boxplot()+
  theme_bw() +
  labs(x = 'Rain Year', y = 'log(Total ticks)\n2016 - 2019',
       title = "Total ticks by rain year") +
  xtitletheme + ytitletheme + rotate_x_text() +
  theme(legend.position = 'bottom', legend.box = 'horizontal')
```



Total ticks ~ temp
```{r}
max <- ggplot(Full_no0, aes(x = mean_maxF, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw()+
  labs(x = 'Mean Max F', y = 'log(Total ticks)\n2016 - 2019')

min <-ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Min F', y = 'log(Total ticks)\n2016 - 2019')


diff <-ggplot(Full_no0, aes(x = temp_diff, y = log(total)))+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Difference in F', y = 'log(Total ticks)\n2016 - 2019')


ggarrange(max, diff + rremove('ylab'), min)

(overlap <- ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
  stat_smooth(data = Full_no0,aes(x = mean_maxF,ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3, color = 'darkblue')+
  stat_smooth(data = Full_no0,aes(x = mean_minF,ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3, color = 'blue')+
  theme_bw()+
  labs(x = 'Mean Min/Max in F', y = 'log(Total ticks)\n2016 - 2019'))

#### look at site

(max_site <- ggplot(Full_no0, aes(x = mean_maxF, y = log(total)))+
  facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw()+
  labs(x = 'Mean Max F', y = 'log(Total ticks)\n2016 - 2019'))

(min_site <-ggplot(Full_no0, aes(x = mean_minF, y = log(total)))+
    facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Min F', y = 'log(Total ticks)\n2016 - 2019'))


(diff_site <-ggplot(Full_no0, aes(x = temp_diff, y = log(total)))+
    facet_wrap(~site)+
  stat_smooth(aes(ymin = ifelse(..ymin.. < 0, 0, ..ymin..)), alpha = .3)+
  theme_bw() +
  labs(x = 'Mean Difference in F', y = 'log(Total ticks)\n2016 - 2019'))

ggarrange(max_site, diff_site + rremove('ylab'), min_site)

# overlap 3 sites in 1 plot for DIFF temp


```


Total ticks ~ precipitation
```{r}
# ggplot(Full_no0, aes(y = log(precip_mm), x = site))+
#   geom_histogram(stat = "identity")+
#   geom_jitter(alpha = .3)+
#   theme_bw()+
#   facet_wrap(~year)+
#   labs(y = 'Precipiation (mm)', x = 'Site')+
#   scale_fill_manual(values =c('brown1','tan1','lightgoldenrod1'))
```

## Stats

Mann-Whitney U/Wilcoxon Rank Sum
- compare medians of arid/mesic (site)
 **wilcox  p = 4.898e-05**
- compare medians of open/total (plot)
  **wilcox p = 5.649e-07**
  
### Median tick counts vary by treatment type for plot and site
+ plot may have a larger difference of tick medians than site
```{r}
site_subset <- Full[which(Full$site == "Arid" | Full$site == "Mesic"),]
plot_subset <- Full[which(Full$plot == "Open" | Full$plot == "Total"),]

ggplot(site_subset, aes(x = total)) + # dataset = hearing, variable of interest = hearing
  geom_histogram() + # geom to make histogram
  facet_wrap(~ site) + # make separate "facets" or grids of your group data by mammal type
  theme_bw()

ggplot(plot_subset, aes(x = total)) + # dataset = hearing, variable of interest = hearing
  geom_histogram() + # geom to make histogram
  facet_wrap(~ plot) + # make separate "facets" or grids of your group data by mammal type
  theme_bw()


ggplot(site_subset, aes(x = site, y = total)) +
  geom_boxplot(fill = "grey89") + # geom to make boxplot, fill the boxplot with the color "grey89"
  theme_classic()

ggplot(plot_subset, aes(x = plot, y = total)) +
  geom_boxplot(fill = "grey89") + # geom to make boxplot, fill the boxplot with the color "grey89"
  theme_classic()

leveneTest(site_subset$total ~ site_subset$site) # df 1, F value = 9.3163, p = 0.002337, < 0.05, variances are different
leveneTest(plot_subset$total ~ plot_subset$plot) # df 1, F value = 26.924, p = 2.599e-07, < 0.05, variances are different


## wilcox (but may data may not meet these assumptions)

wilcox.test(total ~ site, data = site_subset) # p = 4.898e-05
wilcox.test(total ~ plot, data = plot_subset) # p = 5.649e-07
```


ANOVA
- site
  Fvalue = 41.62 
  p < 2e-16 *one site type mean is significantly different*
  
- Plot
  Fvalue = 11.27 
  p = 1.61e-05 *one plot type mean is significantly different*
  
  
Tukey
- site
  Arid, Mesic, Intermediate are all different

-plot
  Open is different than Partial and Total
```{r}
site1 <- aov(log(total)~ site, data = Full_no0)
plot(site1)

# check residuals
res = site1$residuals
shapiro.test(res) # p << 0.05 so data not normal
#hist(res)
leveneTest(total ~ site, 
        data = Full_no0) # p << 0.05 so variances are not equal
summary(site1)
# Fvalue = 41.62 
# p < 2e-16 , reject H0 and conclude that 1 site is significantly different

#Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(site1)
# let's do this in a visual way
Post_hoc <-glht(site1, linfct = mcp(site = 'Tukey'))
summary(Post_hoc)

confint(Post_hoc) # pairwise test
cld(Post_hoc) # arid & mesic = a ; intermediate = b , so intermediate is different than arid & mesic
plot(Post_hoc) # plot 95% CIs

#### ^^ means ot total tick is significantly different in intermediate than mesic and arid

ggplot(data = Full_no0,
       aes(x = site, y = total,
           fill = site))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('brown1','tan1','lightgoldenrod1'))+
  theme_classic()+
  labs(x = 'Site', y = 'log(Total Ticks)', title = 'Total ticks by site')+
  #theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 3.7,
           size = 9) +
  annotate('text', label = 'b', 
           x = 2, y = 8.5,
           size = 9) +
    annotate('text', label = 'a', 
           x = 3, y = 4.1,
           size = 9)



## plot 

Full_no0$plot <- as.factor(Full_no0$plot)
plot1 <- aov(log(total)~ plot, data = Full_no0)
plot(plot1)

# check residuals
res = plot1$residuals
shapiro.test(res) # p << 0.05 so data not normal
#hist(res)
leveneTest(total ~ plot, 
        data = Full_no0) # p << 0.05 so variances are not equal
summary(plot1)
# Fvalue = 11.27 
# p = 1.61e-05 , reject H0 and conclude that 1 site is significantly different

#Tukey-Kramer test assumes that I've already ran an ANOVA & rejected my H0 (all means are equal across groups)
TukeyHSD(plot1)
# let's do this in a visual way
Post_hoc <-glht(plot1, linfct = mcp(plot = 'Tukey'))
summary(Post_hoc)

confint(Post_hoc) # pairwise test
cld(Post_hoc) # arid & mesic = a ; intermediate = b , so intermediate is different than arid & mesic
plot(Post_hoc) # plot 95% CIs

#### ^^ means ot total tick is significantly different in intermediate than mesic and arid

ggplot(data = Full_no0,
       aes(x = plot, y = total,
           fill = plot))+
  geom_bar(stat = 'summary', fun.y = 'mean')+
  stat_summary(geom = 'errorbar', fun.data = mean_se, width = .4)+
  scale_fill_manual(values = c('darkseagreen1','darkseagreen3','darkseagreen4'))+
  theme_classic()+
  labs(x = 'Site', y = 'log(Total Ticks)', title = 'Total ticks by plot')+
  #theme(plot.title = element_text(face = 'bold', hjust = 0.5)) +
  annotate('text', label = 'a', 
           x = 1, y = 4.3,
           size = 9) +
  annotate('text', label = 'b', 
           x = 2, y = 5.9,
           size = 9) +
    annotate('text', label = 'b', 
           x = 3, y = 7.1,
           size = 9)


```

GLMS

 #**Generalized Linear Models**
#What are good predictors (X) of total tick abundance (Y)?
**Ran *generalized linear models* because data breaks assumptions of normality. Used family poisson because of count data.**
*Object **'tickmodels'** is table of df, AIC, BIC, variables used in glm models*
+ *lowest AIC: all 4 variables (Site, Year, Plot, Month)*
+ *lowest AIC for a single predictor = Year*
  + **Year AIC = 10496**
  + **Site AIC = 10665**
  + **Plot AIC = 10871**
models : tick1 - tick7 

```{r GLM for tick data}
tick1 <- glm(total ~ rain_year, 
               data = Full_no0,
             family = 'poisson')
tick2 <- glm(total ~ site, 
               data = Full_no0,
              family = 'poisson')
tick3 <- glm(total ~ plot, 
               data = Full_no0,
              family = 'poisson')
tick4 <- glm(total ~ month, 
               data = Full_no0,
              family = 'poisson')
tick5 <- glm(total ~ site + rain_year, 
               data = Full_no0,
              family = 'poisson')
tick6 <- glm(total ~ site + rain_year + plot, 
               data = Full_no0,
              family = 'poisson')
tick7 <- glm(total ~ site + rain_year + plot + month, 
               data = Full_no0,
              family = 'poisson')
# calculate AIC of each model
tickresults <- AIC(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7)
# add other metrics to table
tickmodels <- list(tick1, tick2, tick3,
                    tick4, tick5, tick6,
                   tick7) # keep models in the same order as they were created
# BIC places higher penalty on extra parameter
tickresults$BIC <- sapply(tickmodels, BIC) # add column for BIC
tick_model_summary <- lapply(tickmodels, summary)
# make for loop to extract r^2 and r^adj from each model
for(i in 1:length(tickmodels)) {
  tickresults$rsq[i] <- tick_model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  tickresults$adj_rsq[i] <- tick_model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again
# add column with variables used in model
tickresults$variables <- c('rain_year','site', 'plot', 'month', 'site + rain_year', 'site + rain_year + plot', 'site + rain_year + plot + month')
# make a nice table!
tickmodels <-kable(tickresults, digits = 2, align = 'c') %>%
  kable_styling()
eighteen <-tickmodels
eighteen
#nineteen <- export_summs(tick1, tick2, tick3, scale = TRUE)
#nineteen
#min(tickresults$AIC)

```


Negative binomial glms

```{r}
# not normal
hist(log(Full_no0$total + 1))
hist(log(Full$total+ 1))

glm.negbin = glm.nb(log(total) ~ site,data = Full_no0)
summary(glm.negbin)

glm.fit = glm(log(total)~site, data = Full_no0,
              family = "poisson")
summary(glm.fit)

glm.quasipoisson = update(glm.fit, family = quasipoisson)
summary(glm.quasipoisson)

# hillary's suggestions
glm1 <- glmer(total ~ mean_minF + plot +  mean_minF*plot + ( 1 | plotID) + (1 | month) + (1 |rain_year),
              data = Full_no0, family = poisson)
summary(glm1)
anova(glm1)
plot(predictorEffects(glm1),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))

```

Full models
```{r}
tick_climate.Full_no0 <- glm.nb(total~site*plot*temp_diff, data=Full_no0)
tick_climate.Full_no0_nointeraction <- glm.nb(total~site+plot+temp_diff, data=Full_no0) # lowest BIC score
tick_climate.S <- glm.nb(total~site, data=Full_no0)
tick_climate.M <- glm.nb(total~plot, data=Full_no0)
tick_climate.D <- glm.nb(total~temp_diff, data=Full_no0)
tick_climate.SplusM <- glm.nb(total~site+plot, data=Full_no0)
tick_climate.StimesM <- glm.nb(total~site*plot, data=Full_no0)
tick_climate.SplusD <- glm.nb(total~site+temp_diff, data=Full_no0)
tick_climate.StimesD <- glm.nb(total~site*temp_diff, data=Full_no0)
tick_climate.MplusD <- glm.nb(total~plot+temp_diff , data=Full_no0)
tick_climate.MtimesD <- glm.nb(total~plot*temp_diff, data=Full_no0)
tick_climate.SplusMD <- glm.nb(total~site+plot*temp_diff, data=Full_no0)
tick_climate.MplusSD <- glm.nb(total~plot+site*temp_diff, data=Full_no0)
tick_climate.DplusSM <- glm.nb(total~site*plot+temp_diff, data=Full_no0)
AIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
BIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
temp_diff <- BIC(tick_climate.Full_no0,tick_climate.Full_no0_nointeraction,tick_climate.S,tick_climate.M,tick_climate.D,
    tick_climate.SplusM,tick_climate.StimesM, tick_climate.SplusD, tick_climate.StimesD, 
    tick_climate.MplusD, tick_climate.MtimesD, tick_climate.SplusMD, tick_climate.MplusSD,
    tick_climate.DplusSM)
min_dif <-min(temp_diff$BIC)
plot(predictorEffects(tick_climate.Full_no0, ~ plot),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
plot(predictorEffects(tick_climate.Full_no0, ~ site),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
plot(predictorEffects(tick_climate.Full_no0, ~ temp_diff),
     lines=list(multiline=TRUE),
     lattice=list(key.args=list(columns=2)),
     confint=list(style="bars"))
```


Hurdle model
```{r}
## part 1 of 0s
# zero inflated poisson (ZIP)
fldown <- formula(total ~ site + plot | site + plot) # what explains the load that's not 0 | explanatory variables if something is present or not
ZIP_round_down <- zeroinfl(fldown, dist = 'poisson',
                           link = 'logit',
                           data = Full)
summary(ZIP_round_down) # site and plot are signficant!
# zero inflated negative binomial (ZINB)
fldown <- formula(total ~ site + plot | site)
ZINB_round_down <- zeroinfl(fldown, dist = 'negbin', # changed distribution
                            link = 'logit',
                            data = Full)
summary(ZINB_round_down) # site and plot are significant!
# likeligood ratio test to compare ZIP and ZINB
# test the two models
lrtest(ZIP_round_down, ZINB_round_down) # ZINB p < 2.2e-16, so go with negative binomial
## part two non-0s
# Hurdle Models
# ZIP
H1downA <- hurdle(fldown, dist = "poisson", link = "logit",data = Full)
# ZINB
H1downB <- hurdle(fldown, dist = "negbin", link = "logit",data = Full)
lrtest(H1downA,H1downB) # Negative binomial version much more significant, p < 2.2e-16
summary(H1downB)
```

Nonlinear least squares (logistic growth fitting)
```{r}
library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car) #for plotting
```

###Part 1. Nonlinear least squares (logistic growth fitting)

a. Load dataset Full.csv. Create a scatterplot.

```{r get_data}
Full <- read_csv("Full.csv")
```

b. Look at it

```{r cell_graph} 
ggplot(Full_no0, aes(x = temp_diff, y = total)) + geom_point()
```

Recall logistic growth equation: 

N(t) = A/(1+Be^-rt)

# Estimate for K/A: ~ 3700
# Estimate for N0: ~ 2000
# Estimate for B: ~ 0.85

BEstimate <- (3700 - 2000)/2000 # BEstimate = 0.85

c. Create a subset of the data that you think is just in the 'exponential growth phase.' Take the natural log of the count and create a basic scatterplot of temp_diff v ln(counts). 

```{r}
Graph2 <- plot(Full_no0$temp_diff,log(Full_no0$total))
# Create a basic scatterplot of temp_diff versus ln(totals)

fit <- lm(log(total + 1) ~ temp_diff, data = Full_no0)

hist(fit$residuals)
```

Then find the slope of that line to get 'r' (growth rate constant)

```{r}
r_est <- lm(log(Full_no0$total) ~ Full_no0$temp_diff) # R ~ 0.1035
```


d. Nonlinear least squares 

e. Model fitting 

```{r}
cell_fit <- nls(total ~ A/(1+B*exp(-r*temp_diff)), 
                start = list(A = 2000, B = 0.02088, r = 0.44187), 
                data = Full_no0, trace = TRUE) 
```


Uses the estimates as a start to find the best estimated values (by convergence) for A, B and r

8 iterations for convergence
Estimates, St. Errors and p-values

Parameters:
Estimate Std. Error t value Pr(>|t|)    
A 3.806e+03  5.029e+01   75.67  < 2e-16 ***
B 9.233e-01  5.887e-02   15.68 2.33e-09 ***
r 3.443e-01  3.219e-02   10.70 1.73e-07 ***

f. Create new variables for A, B and r

```{r}
A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]
```

g. Create a new sequence containing a series of temp_diffs over which you'll predict the cell count

```{r}
temp_diff_seq <- seq(0,20, length = 100)
```

h. Using the parameters (A, B, and r) and the temp_diff sequence, predict the cell counts for the logistic growth model

```{r}
cell_pred <- A/(1+B*exp(-r*temp_diff_seq)) # Logistic growth model
```

i. Bind together the temp_diff sequence data and the predictions data into a new data frame

```{r}
pred_df <- data.frame(temp_diff_seq, cell_pred)
```


j. Create a single graph in which you show the original data (as scatterplot points) and the predicted data (as a line graph)

```{r}
ggplot(Full_no0, aes(x = temp_diff, y = total)) + 
  geom_point(colour = "blue", size = 3) + 
  theme_bw() +
  geom_line(data = pred_df, aes(x = temp_diff_seq, y = cell_pred), colour = "orange", size = 1) + 
  xlab("temp_diff (h)") +
  ylab("Cell Count") +
  ggtitle("Bacterial Growth")
```